---
title: "Colocalization Timeseries"
author: "Sven Meinerz"
format: 
  docx:
    reference-doc:  template_class.docx
    toc: true
    toc-depth: 3
    toc-title: Table of Content
editor: visual
execute: 
  echo: false # We get the source code than the output, must of time disabled is best option
  warning: false # Common option, hides warnings; errors will still break the code
  output: asis # as is, rendering will not round numbers etc. Leave the output as it is
fig-dpi: 300
fig-width: 6
fig-height: 8
---

# Setup

```{r}
#| echo: true
pacman::p_load(
 conflicted, tidyverse, here, ggplot2, wrappedtools,
 ggbeeswarm, readxl, ggsignif, patchwork, ggh4x, flextable, TOSTER,
 nlme, broom, DescTools, lme4, merTools, gg4x, lme4
)
conflicts_prefer(
  dplyr::select,
  dplyr::filter,
  ggplot2::mean_cl_boot,
  modelbased::standardize
)
```

# Data Import

Importing data from an excel file. Data is inside the sheet called "rawdata".

```{r}
#| echo: true
rawdata <- read_excel(
  path = here("Data/Coloc_Timeseries.xlsx"),
  sheet = "rawdata"
  )
rawdata <- mutate(
  .data = rawdata,
  Strain = factor(Strain,
                     levels = c("WT", "Mut1", "Mut2"))
)
```

# Graphical Data exploration

## Quick check of data distribution

Here we check distribution of the PCC values and overlay a theoretical, normal distributed curve based on the measured values.

```{r}
#| echo: true

ggplot(rawdata, aes(PCC)) + 
  geom_density(fill = "lightblue", alpha = 0.5) +
  stat_theodensity(distri = "norm", linewidth = 0.5, color = "red") +
  facet_wrap(facets = vars(Strain))

```

Based on these plots we can assume, that the PCC values are not normal distributed, as they deviated from theoretical distribution.

## Exploration for data modelling

Next we look into our data and try to eyeball what an appropriate model for the data might be. First we check for a general linear model.

```{r}
#| echo: true

ggplot(rawdata, aes(Timepoint, PCC, color = Strain)) +
  stat_summary(fun.data = mean_se, position = position_dodge(width = 1)) +
  ylab("PCC\nmean \u00b1 SEM")

ggplot(rawdata, aes(x = Timepoint, y = PCC)) +
  geom_point() + 
  geom_smooth() +
  geom_smooth(method = "lm", color = "red") +
  facet_wrap(facets = vars(Strain))

ggplot(rawdata, aes(Timepoint, PCC)) +
  geom_point() +
  geom_smooth() +
  geom_smooth(method = "lm", color = "red")

ggplot(rawdata, aes(Timepoint, PCC, color = Strain)) +
  geom_point(alpha = 0.25) +
  geom_smooth() +
  scale_y_continuous(name = "PCC\nmean \u00b1 95% CI", limits = c(0, 0.75)) +
  xlab("elapsed time [min]")
```

As we see in the two generated plots, the red line from the linear regression is quiet different from the result from the loess method. This was expected for a number of reasons:\
1. From the experiment performed, we expected at first a rise of the PCC values, than the longer the time goes on that will fall again.\
2. PCC values are hard limited, meaning they always have a value between -1 and 1. It would not be appropriate to use a linear regression to predict values outside of this range.

Let's take a moment and think about our data: We have one measured, dependent variable (PCC) and two independent variables (Strain and Timepoint). Based on this a linear mixed model (LMM) seems to be more appropriate.

# Building the Linear Mixed Model
## Random Intercept Model
```{r}
lmer_out <- lmer(PCC ~ Strain + (1 | Timepoint), data = rawdata)
lmer_out
(Anova(lmer_out))
```

```{r}
(lmer_out_param <- model_parameters(lmer_out, group_level = TRUE))
intercept_all <- lmer_out_param$Coefficient[1]
slope_Ang <- lmer_out_param$Coefficient[2]
slope_What <- lmer_out_param$Coefficient[3]
plotdata <- tibble(
  Timepoint = lmer_out_param$Level[-(1:3)],
  intercept = lmer_out_param$Coefficient[-(1:3)] +
    intercept_all
)

```

The `echo: false` option disables the printing of code (only output is displayed).
