---
title: "Demo Exam Statistics"
author: Your Name
format: docx
editor: visual
markdown: 
  wrap: 60
---

# Any packages needed? put them here!

```{r}
pacman::p_load(conflicted, tidyverse, broom, car)
```

Please give your answers underneath the question, referencing question numbers. Please answer outside code chunks or format text answers as comments.

# Q1: Please generate the following sequences of numbers:

## q1a: Numbers from 100 to 199

```{r}
#q1a<-c(100,101,102...)
seq(100, 199, 1)
```

## q1b: 100 random data from a uniform distribution between 100 and 199

```{r}
runif(100, 100, 199)
```

## q1c: 100 random data from a normal distribution with mean=100 and variance=10

```{r}
rnorm(100, 100, sqrt(10))
```

## q1d: 100 random data from a log normal distribution with log(mean)=5 and log(sd)=0.5

```{r}
rlnorm(100, 5, 0.5)
```

## q1e: 100 random data from a factor with levels 1 and 2 in a ratio of 7:3.

```{r}
x <- sample(x = c(1, 2), size = 100, replace = TRUE, 
            prob = c(0.7, 0.3)) |> 
  as.factor()
```

# Q2: Please name statistical tests or class of analyses for the following analytical problems:

## q2a: Test for normal distribution

Shapiro-Wilk, Kolmogorov-Smirnov test ...

## q2b: Multiple comparison for means between 3 or more groups for a Gaussian measure

ANOVA, lm(), post-hoc like pairwise.t.test, Dunnet, Tukey ...

# Q3: Assign the dataset mtcars to a new tibble

```{r}
 new_tibble <- as_tibble(mtcars)
class(new_tibble)
```

# Q4: For the columns in Q3, please state if the data most likely represent Gaussian measures, ordinal measures, or categories

am, cyl... are likely factors

```{r}
fact_vars <- c("am","cyl")
```

# Q5: For all numerical variables, test for Normal distribution.

```{r}
```

# Q6

## q6a: For all data assumed to follow a normal distribution, test for group differences between the various numbers of cylinders, post-hoc test optional

## q6b: Save descriptive statistics for groups and p-value for a group effect in one or several result tables

```{r}
```

## q6c: Export the result table(s) in an Excel file

```{r}
write_excel(df, "name.xlsx")
```

# Q7: Test regression between independent variable wt and dependent variable mpg

## q7a: report the regression parameters

```{r}
lm_out <- lm(mpg ~ wt, data = new_tibble)
intercept <- lm_out$coefficients[1]
slope <- lm_out$coefficients[2]
```

## q7b: report significance for regression parameters

```{r}
anova_out <- anova(lm_out)
p <- pluck(anova_out$`Pr(>F)`)
p <- p$p.value # rather use pluck()
```

# Q8: Test the relation between dependent variable am and independent variables wt and mpg

## q8a: report the model parameters

```{r}
lm_out <- lm(am ~ wt + mpg, data = new_tibble)
intercept <- lm_out$coefficients[1]
slope_wt <- lm_out$coefficients[2]
slope_mpg <- lm_out$coefficients[3]
```

## q8b: report significance for model parameters

```{r}
anova_out <- Anova(lm_out)
p <- tidy(anova_out)
p <- p$p.value
```

Good luck!
