---
title: "Demo Exam Statistics"
author: Jan Itor
format: docx
editor: visual
markdown: 
  wrap: 60
---

# Any packages needed? put them here!

```{r}
pacman::p_load(conflicted,tidyverse,wrappedtools,writexl, car, broom)
```

Please give your answers underneath the question, referencing question numbers. Please answer outside code chunks or format text answers as comments.

# Q1: Please generate the following sequences of numbers:

## q1a: Numbers from 100 to 199

```{r}
#q1a<-c(100,101,102...)
q1a2 <- 100:199
q1a3 <- seq(100,199,1)
```

## q1b: 100 random data from a uniform distribution between 100 and 199

```{r}
q1b <- runif(n = 100,min = 100,max = 199)
```

## q1c: 100 random data from a normal distribution with mean=100 and variance=10

```{r}
q1c <- rnorm(n = 100,mean = 100,sd = sqrt(10))
```

## q1d: 100 random data from a log normal distribution with log(mean)=5 and log(sd)=0.5

```{r}
q1d <- rlnorm(n = 100,meanlog = 5,sdlog = 0.5)
```

## q1e: 100 random data from a factor with levels 1 and 2 in a ratio of 7:3.

```{r}
q1e <- sample(x = c(1,2),size = 100,replace = TRUE,
              prob = c(0.7,0.3)) |> as.factor()
set.seed(321)
q1e <- sample(x = factor(c(1,2)),size = 100,replace = TRUE,
              prob = c(0.7,0.3))
```

# Q2: Please name statistical tests or class of analyses for the following analytical problems:

## q2a: Test for normal distribution

Shapiro-Wilk test, Kolmogorov-Smirnov test; ks.test() ..

## q2b: Multiple comparison for means between 3 or more groups for a Gaussian measure

ANOVA, lm(), post-hoc tests like pairwise.t.test, Dunnet, Tukey...

# Q3: Assign the dataset mtcars to a new tibble

```{r}
newby <- as_tibble(mtcars)
class(newby)
```

# Q4: For the columns in Q3, please state if the data most likely represent Gaussian measures, ordinal measures, or categories

I treat cyl, cs, and am as factors.

```{r}
str(newby)
fact_vars <- c("cyl","vs","am","gear")
ord_vars <- c("carb")
gauss_vars <- c("mpg","disp","hp","drat","wt","qsec")
```

# Q5: For all numerical variables, test for Normal distribution.

```{r}
summarize(newby,
          across(where(is.numeric),
                 ~shapiro.test(.x)$p.value)) |>
  pivot_longer(everything()) |>
  mutate(Gaussian=case_when(value>=0.05~"yes",
                            .default="no"))
gauss_vars <- ColSeeker(newby,namepattern = c("t","mpg","q"))
ord_vars <- ColSeeker(newby, namepattern = c("cyl","dis","hp"))
```

# Q6

## q6a: For all data assumed to follow a normal distribution, test for group differences between the various numbers of cylinders

## q6b: Save descriptive statistics for groups and p-value for a group effect in one or several result tables

```{r}
# approach 1 with single function:
all_out <-
  compare_n_numvars(.data = newby,
                    dep_vars = gauss_vars$names,
                    indep_var = "cyl",gaussian = TRUE)
q6ab1 <- 
  all_out$results |>
  select(Variable,contains(" fn"), multivar_p) 
# approach 2 with summarize(acros...) in 2 steps
q6a <- newby |>
  group_by(cyl) |>
  summarize(across(all_of(gauss_vars$names),
                   ~meansd(.x,range = T))) |>
  pivot_longer(-cyl,names_to = "Variable",
               values_to = "DescStats") |>
  pivot_wider(names_from = cyl,values_from = DescStats)
q6b <- newby |>
  summarize(across(all_of(gauss_vars$names),
                   ~lm(.x~cyl) |>
                     anova() |>
                     slice(1) |>
                     pull("Pr(>F)") |>
                     formatP(5,mark = T))) |>
  pivot_longer(everything(), names_to="Variable",
               values_to="p-value")
q6ab2 <- full_join(q6a,q6b)
# approach 3 with a loop, here just for posthoc but can be extend
# loop for pairwise t-tests
posthoc_out <- tibble(Variable=NA_character_,
                      p4vs.6=NA_character_,
                      p4vs.8=NA_character_,
                      p6vs.8=NA_character_,
                      .rows = 0)
for(var_i in gauss_vars$names){
  t_out <- pairwise_t_test(newby[[var_i]],newby$cyl,
                           adjmethod="fdr")$p.value |>
    formatP()
  posthoc_out <- posthoc_out |>
    add_row(Variable=var_i,
            p4vs.6=t_out[1,1],
            p4vs.8=t_out[2,1],
            p6vs.8=t_out[2,2])
}
```

## q6c: Export the result table(s) in an Excel file

```{r}
write_xlsx(q6ab1,"Q6ab.xlsx")
write_xlsx(list(desc=q6a,anova=q6b,posthoc=posthoc_out),
           "Q6.xlsx")
```

# Q7: Test regression between independent variable wt and dependent variable mpg

## q7a: report the regression parameters

```{r}
lmq7 <- lm(mpg~wt,data = newby)
cat("Regression parameters: \n", "Intercept: ", lmq7$coefficients[1],
    "Slope: ", lmq7$coefficients[2])
```

## q7b: report significance for regression parameters

```{r}
anova(lmq7) |>
  pluck("Pr(>F)") |>
  na.omit() |>
  formatP(5,mark = T)
```

# Q8: Test the relation between dependent variable am and independent variables wt and mpg

## q8a: report the model parameters

```{r}
lmq8 <- glm(am~wt+mpg,data = newby,
            family=binomial())
cat("Regression parameters: \n", "Intercept: ", lmq8$coefficients[1],
    " OR wt: ", exp(lmq8$coefficients[2]),
    "\n OR mpg: ", exp(lmq8$coefficients[3]))
```

## q8b: report significance for model parameters

```{r}
car::Anova(lmq8) |>
  tidy() |> 
  mutate(p.value = formatP(p.value, ndigits = 5,mark = T))
```

Good luck!
