---
title: "My first document"
author: "Sven Meinerz"
format: 
  docx:
    reference-doc:  template_class.docx
    toc: true
    toc-depth: 3
    toc-title: Table of Content
editor: visual
execute: 
  echo: false # We get the source code than the output, must of time disabled is best option
  warning: false # Common option, hides warnings; errors will still break the code
  output: asis # as is, rendering will not round numbers etc. Leave the output as it is
fig-dpi: 300
fig-width: 6
fig-height: 8
---

\newpage

This report was created at `r Sys.time()`. Less accurately, this report was created at `r Sys.Date()`. Even worse, some say it was `r format(Sys.time(), "%Y-%m-%d %H:%M")`.

YAML

# Quarto

***Quarto*** enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

First chunk should contain all the setup, like packages

```{r}
pacman::p_load(tidyverse, flextable)
set_flextable_defaults(
  big.mark = " ",
  font.size = 7,
  theme_fun = theme_zebra,
  padding.bottom = 1,
  padding.top = 1,
  padding.left = 3,
  padding.right = 4
)
theme_set(theme_bw())
```

This is what a table could look like:

```{r}
penguins |> 
  flextable() |> 
  # theme_zebra() |> 
  set_table_properties(width = 1,layout = 'autofit')
```

Let's do some fancy table things:

```{r}
#| echo: true
penguins |> 
  flextable() |>
  separate_header() |>
  bg(~sex == "female", j = c(1,7), bg = "pink") |>
  bold(~island == "Torgersen", 2) |>
  set_table_properties(width = 1,layout = 'autofit')
```

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
