---
title: "HDLox2024"
author: "Dr. Jan Itor"
format: 
  docx:
    reference-doc: template2026_2.docx
    toc: true
    toc-depth: 4
    output-file: HDLoxReport202609xx.docx 
execute:
  echo: false
  warning: false
  output: asis
fig-dpi: 300
fig-width: 6
fig-height: 4
editor: visual
---

\newpage

# *Background*

*Does HDLox differ between subject with/without PAD? Can it improve prediction/diagnosis of PAD?*

# *Methods*

# *Analysis*

## *Setup and import*

```{r}
pacman::p_load(
  conflicted, tidyverse, wrappedtools,
  broom, flextable, readxl, here,
  car, multcomp, pROC,
  ggbeeswarm, GGally, ggsignif,
  polyglotr
)
conflicts_prefer(
  dplyr::filter,
  dplyr::select
)
set_flextable_defaults(
  big.mark = " ",
  font.size = 9,
  theme_fun = theme_vanilla,
  padding.bottom = 1,
  padding.top = 3,
  padding.left = 3,
  padding.right = 4
)
set_theme(theme_light())
pad_colors <- c("darkolivegreen", "darkorange")
```

```{r}
#| echo: false
rawdata <-
  read_excel(here("data/DatensatzHDLox.xlsx")#,
             # range = cell_cols("A:AB")
  ) |>
  select(`Patienten-ID`:`Smoking (Nein= 0, Ja=1)`) |>
  select(
    -Geburtsdatum,
  ) |>
  rename(
    `LP(a) nmol/l` = "LP(a) (Lipoprotein a, nmol/l) cut off 120nmol/l",
    CAD = "Angiographically confirmed CAD (1=ja, 0=nein), 0 bedeudet angiogr. Ausgeschlossen",
    PAD = "Peripheral/Carotid artery disease (pAVK/cAVK) (0=nein, 1=ja)",
    Adipositas = "Adipositas (cut off BMI ≥ 30 kg/m2, 0-keine Adipositas, 1-Adipositas Grad 1 oder höher)"
  ) |>
  rename_with(
    \(x) str_replace_all(
      x, c(
        " \\( *[01]=.*" = "", # variations of ja nein male 
        # " \\( *1=.*" = "",
        " \\(Nein=.*" = "",
        " \\( *[uU].*" = "", # Umrechnung
        "NT " = "NT",
        "\\.{3}\\d+" = "", # added ... for repeated names
        " in" = "", # prefix for some m. units
        ", .+" = "",
        " (.+/.+)" = " [\\1]"
      )
    ) |>
      str_squish()
  ) |>
  mutate(
    `Patienten-ID` = as.character(`Patienten-ID`),
    Gender = factor(Gender,
                    levels = c(1, 2),
                    labels = c("male", "female")
    ),
    CAD = replace_na(CAD, 0),
    `LP(a) [nmol/l]` =
      str_replace_all(`LP(a) [nmol/l]`, c(
        "<" = "",
        "," = "."
      )) |>
      as.numeric() |>
      round(1),
    across(
      where( \(x) (length(unique(na.omit(x))) == 2 &
                 all(unique(na.omit(x)) %in% c(0, 1))
      )),
      # c(`Healthy subjects`,
      #        CAD,PAD, `Diabetes mell.`,
      #        Hypertension, Smoking, Microalbuminuria,
      #        Adipositas),
      .fns = ~ factor(
        .x,
        levels = c(0, 1),
        labels = c("no", "yes")
      )
    ),
    Größe = case_when(Größe > 3 ~ Größe / 100,
                      .default = Größe
    ),
    BMI = Gewicht / Größe^2
  ) |>
  # filter_out(is.na(HDLox))
  drop_na(HDLox, PAD)

german_names <- ColSeeker(
  namepattern = c("Pat", "G[re]", "Adip", "Chol")
)
colnames(rawdata)[german_names$index] <-
  polyglotr::google_translate(german_names$names,
                              target_language = "en",
                              source_language = "de"
  )
cn()
# n_obs <- nrow(rawdata)
# n_complete <- nrow(drop_na(rawdata))
```

\newpage

## *Test for Normal distribution*

```{r}
#| fig.height: 3
gausstest <-
  rawdata |>
  group_by(`Healthy subjects`) |>
  summarize(across(
    where(is.numeric),
    list(
      "norm" = \(x) ksnormal(x,
                          lillie = FALSE
      ),
      "lognorm" = \(x) ksnormal(log(x),
                             lillie = FALSE
      )
    )
  )) |>
  pivot_longer(-`Healthy subjects`,
               names_to = c("Variable", ".value"),
               names_sep = "_"
  ) |>
  arrange(Variable, `Healthy subjects`) |>
  pivot_wider(
    names_from = `Healthy subjects`,
    values_from = contains("norm"),
    names_glue = "healthy:{`Healthy subjects`} {.value}"
  ) |>
  mutate(across(
    contains("Healthy"),
    ~ formatP(.x,
              ndigits = 5, mark = TRUE
    )
  ))
flextable(gausstest) |>
  bg(~ str_detect(`healthy:no norm`, "n.s."),
     j = 2, bg = "lightgreen"
  ) |>
  bg(~ str_detect(`healthy:yes norm`, "n.s."),
     j = 3, bg = "lightgreen"
  ) |>
  bg(~ str_detect(`healthy:no lognorm`, "n.s."),
     j = 4, bg = "lightgreen"
  ) |>
  bg(~ str_detect(`healthy:yes lognorm`, "n.s."),
     j = 5, bg = "lightgreen"
  ) |>
  set_table_properties(
    width = 1,
    layout = "autofit"
  )
cat("<br>\n\n")
test_vars <- ColSeeker(rawdata,
                       varclass = "numeric",
                       exclude = c("Age")
)
log_vars <- ColSeeker(namepattern = c(
  "DL ", "Chol", "Tri",
  "BNP", "hsC", "Alb"
))
for (var_i in test_vars$names) {
  plot_out <-
    ggplot(rawdata, aes(
      x = .data[[var_i]],
      fill = `Healthy subjects`
    )) +
    geom_density(alpha = .5) +
    ggtitle(paste(
      "Control", gausstest |>
        filter(Variable == var_i) |>
        pull(`healthy:yes norm`),
      "Patient", gausstest |>
        filter(Variable == var_i) |>
        pull(`healthy:no norm`),
      collapse = "\n"
    )) +
    scale_fill_manual(values = pad_colors)
  print(plot_out)
  cat("<br>\n\n")
  if (var_i %in% log_vars$names) {
    plot_out2 <-
      ggplot(rawdata, aes(
        x = .data[[var_i]],
        fill = `Healthy subjects`
      )) +
      geom_density(alpha = .5) +
      scale_x_log10(
        breaks = logrange_15,
        minor_breaks = logrange_123456789
      ) +
      ggtitle(paste(
        "Control", gausstest |>
          filter(Variable == var_i) |>
          pull(`healthy:yes lognorm`),
        "Patient", gausstest |>
          filter(Variable == var_i) |>
          pull(`healthy:no lognorm`),
        collapse = "\n"
      )) +
      scale_fill_manual(values = pad_colors)
    print(plot_out2)
  }
  cat("<br>\n\n")
}
```

```{r}
log_vars <- ColSeeker(namepattern = c(
  "DL ", "Chol", "Tri",
  "BNP", "hsC"
))
rawdata <-
  rawdata |>
  mutate(across(all_of(log_vars$names),
                ~ log(.x + 0.001),
                .names = "{.col} (ln)"
  ))
```

*When a lognormal distribution is assumed, variables will be log-transformed and treated as gaussian AND will go into ordvars as well.*

```{r}
fact_vars <- ColSeeker(rawdata,
                       varclass = "factor",
                       exclude = "PAD"
)
gauss_vars <- ColSeeker(rawdata,
                        varclass = "numeric",
                        namepattern = c(
                          "\\(ln\\)", "IMT",
                          "^Si", "W", "BMI"
                        )
)
ord_vars <- ColSeeker(rawdata,
                      varclass = "numeric",
                      namepattern = c(
                        "DL", "Tri", "Chol", "Albu",
                        "LP", "hsC", "PLA",
                        "BNP", "GFR",
                        "HbA", "Age"
                      ),
                      exclude = "\\(ln\\)"
)
```

\newpage

## *Univariate Analysis for PAD*

### *Gaussian Variables*

\newpage

### *Ordinal Variables*

\newpage

### *Factor Variables*

\newpage

## *Correlations*

## *Prediction analysis*

### *All univariable models*

\newpage

### *Multivariable model*
