---
title: "HDLox2024"
author: "Dr. Jan Itor"
format: 
  docx:
    reference-doc: template2026_2.docx
    toc: true
    toc-depth: 4
    output-file: HDLoxReport202609xx.docx 
  typst:
    toc: true
    lof: true
    lot: true
    keep-typ: true
execute:
  echo: false
  warning: false
  output: asis
fig-dpi: 300
fig-width: 6
fig-height: 4
editor: visual
---

\newpage

# Background

Does HDLox differ between subject with/without PAD? Can it improve prediction/diagnosis of PAD?

# Methods

# Analysis

## Setup and import

```{r}
pacman::p_load(
  conflicted, tidyverse, wrappedtools,
  broom, flextable, readxl, here,
  car, multcomp, pROC,
  ggbeeswarm, GGally, ggsignif,
  polyglotr
)
conflicts_prefer(
  dplyr::filter,
  dplyr::select
)
set_flextable_defaults(
  big.mark = " ",
  font.size = 9,
  theme_fun = theme_vanilla,
  padding.bottom = 1,
  padding.top = 3,
  padding.left = 3,
  padding.right = 4
)
set_theme(theme_light())
pad_colors <- c("darkolivegreen", "darkorange")

internal_version <- TRUE
```

```{r}
#| echo: false
rawdata <-
  read_excel(here("data/DatensatzHDLox.xlsx")#,
             # range = cell_cols("A:AB")
  ) |>
  select(`Patienten-ID`:`Smoking (Nein= 0, Ja=1)`) |>
  select(
    -Geburtsdatum,
  ) |>
  rename(
    `LP(a) nmol/l` = "LP(a) (Lipoprotein a, nmol/l) cut off 120nmol/l",
    CAD = "Angiographically confirmed CAD (1=ja, 0=nein), 0 bedeudet angiogr. Ausgeschlossen",
    PAD = "Peripheral/Carotid artery disease (pAVK/cAVK) (0=nein, 1=ja)",
    Adipositas = "Adipositas (cut off BMI ≥ 30 kg/m2, 0-keine Adipositas, 1-Adipositas Grad 1 oder höher)"
  ) |>
  rename_with(
    \(x) str_replace_all(
      x, c(
        " \\( *[01]=.*" = "", # variations of ja nein male 
        # " \\( *1=.*" = "",
        " \\(Nein=.*" = "",
        " \\( *[uU].*" = "", # Umrechnung
        "NT " = "NT",
        "\\.{3}\\d+" = "", # added ... for repeated names
        " in" = "", # prefix for some m. units
        ", .+" = "",
        " (.+/.+)" = " [\\1]"
      )
    ) |>
      str_squish()
  ) |>
  mutate(
    `Patienten-ID` = as.character(`Patienten-ID`),
    Gender = factor(Gender,
                    levels = c(1, 2),
                    labels = c("male", "female")
    ),
    CAD = replace_na(CAD, 0),
    `LP(a) [nmol/l]` =
      str_replace_all(`LP(a) [nmol/l]`, c(
        "<" = "",
        "," = "."
      )) |>
      as.numeric() |>
      round(1),
    across(
      where( \(x) (length(unique(na.omit(x))) == 2 &
                 all(unique(na.omit(x)) %in% c(0, 1))
      )),
      # c(`Healthy subjects`,
      #        CAD,PAD, `Diabetes mell.`,
      #        Hypertension, Smoking, Microalbuminuria,
      #        Adipositas),
      .fns = ~ factor(
        .x,
        levels = c(0, 1),
        labels = c("no", "yes")
      )
    ),
    Größe = case_when(Größe > 3 ~ Größe / 100,
                      .default = Größe
    ),
    BMI = Gewicht / Größe^2
  ) |>
  # filter_out(is.na(HDLox))
  drop_na(HDLox, PAD)

german_names <- ColSeeker(
  namepattern = c("Pat", "G[re]", "Adip", "Chol")
)
colnames(rawdata)[german_names$index] <-
  polyglotr::google_translate(german_names$names,
                              target_language = "en",
                              source_language = "de"
  )
if (internal_version) {
  cn()
}
  # n_obs <- nrow(rawdata)
# n_complete <- nrow(drop_na(rawdata))
```

\newpage

## Test for Normal distribution

```{r}
#| fig.height: 3
gausstest <-
  rawdata |>
  group_by(`Healthy subjects`) |>
  summarize(across(
    where(is.numeric),
    list(
      "norm" = \(x) ksnormal(x,
                          lillie = FALSE
      ),
      "lognorm" = \(x) ksnormal(log(x),
                             lillie = FALSE
      )
    )
  )) |>
  pivot_longer(-`Healthy subjects`,
               names_to = c("Variable", ".value"),
               names_sep = "_"
  ) |>
  arrange(Variable, `Healthy subjects`) |>
  pivot_wider(
    names_from = `Healthy subjects`,
    values_from = contains("norm"),
    names_glue = "healthy:{`Healthy subjects`} {.value}"
  ) |>
  mutate(across(
    contains("Healthy"),
    ~ formatP(.x,
              ndigits = 5, mark = TRUE
    )
  ))
if (internal_version) {
  flextable(gausstest) |>
    bg(~ str_detect(`healthy:no norm`, "n.s."),
       j = 2, bg = "lightgreen"
    ) |>
    bg(~ str_detect(`healthy:yes norm`, "n.s."),
       j = 3, bg = "lightgreen"
    ) |>
    bg(~ str_detect(`healthy:no lognorm`, "n.s."),
       j = 4, bg = "lightgreen"
    ) |>
    bg(~ str_detect(`healthy:yes lognorm`, "n.s."),
       j = 5, bg = "lightgreen"
    ) |>
    set_table_properties(
      width = 1,
      layout = "autofit"
    )
  cat("<br>\n\n")
}
test_vars <- ColSeeker(rawdata,
                         varclass = "numeric",
                         exclude = c("Age")
  )
log_vars <- ColSeeker(namepattern = c(
  "DL ", "Chol", "Tri",
  "BNP", "hsC", "Alb"
))
if (internal_version) {
  for (var_i in test_vars$names) {
  plot_out <-
    ggplot(rawdata, aes(
      x = .data[[var_i]],
      fill = `Healthy subjects`
    )) +
    geom_density(alpha = .5) +
    ggtitle(paste(
      "Control", gausstest |>
        filter(Variable == var_i) |>
        pull(`healthy:yes norm`),
      "Patient", gausstest |>
        filter(Variable == var_i) |>
        pull(`healthy:no norm`),
      collapse = "\n"
    )) +
    scale_fill_manual(values = pad_colors)
  print(plot_out)
  cat("<br>\n\n")
  if (var_i %in% log_vars$names) {
    plot_out2 <-
      ggplot(rawdata, aes(
        x = .data[[var_i]],
        fill = `Healthy subjects`
      )) +
      geom_density(alpha = .5) +
      scale_x_log10(
        breaks = logrange_15,
        minor_breaks = logrange_123456789
      ) +
      ggtitle(paste(
        "Control", gausstest |>
          filter(Variable == var_i) |>
          pull(`healthy:yes lognorm`),
        "Patient", gausstest |>
          filter(Variable == var_i) |>
          pull(`healthy:no lognorm`),
        collapse = "\n"
      )) +
      scale_fill_manual(values = pad_colors)
    print(plot_out2)
  }
  cat("<br>\n\n")
  }
}
```

```{r}
log_vars <- ColSeeker(namepattern = c(
  "DL ", "Chol", "Tri",
  "BNP", "hsC"
))
rawdata <-
  rawdata |>
  mutate(across(all_of(log_vars$names),
                \(x) log(x + 0.001),
                .names = "{.col} (ln)"
  ))
```

When a lognormal distribution is assumed, variables will be log-transformed and treated as gaussian AND will go into ordvars as well.

```{r}
fact_vars <- ColSeeker(rawdata,
                       varclass = "factor",
                       exclude = "PAD"
)
gauss_vars <- ColSeeker(rawdata,
                        varclass = "numeric",
                        namepattern = c(
                          "\\(ln\\)", "IMT",
                          "^Si", "W", "BMI"
                        )
)
ord_vars <- ColSeeker(rawdata,
                      varclass = "numeric",
                      namepattern = c(
                        "DL", "Tri", "Chol", "Albu",
                        "LP", "hsC", "PLA",
                        "BNP", "GFR",
                        "HbA", "Age"
                      ),
                      exclude = "\\(ln\\)"
)
```

\newpage

## Univariate Analysis for PAD

### Gaussian Variables

```{r}
cat("<br>\n\n")
gauss_out <- compare2numvars(
  rawdata,
  dep_vars = gauss_vars$names,
  indep_var = "PAD",
  gaussian = TRUE,
  add_n = TRUE,
  round_desc = 3
)
gauss_out |>
  select(-desc_all) |>
  flextable() |>
  set_table_properties(
    width = 1,
    layout = "autofit"
  )
cat("<br>\n\n")
```

```{r}
for (var_i in gauss_vars$names) {
  plot_out <-
    ggplot(rawdata, aes(PAD, .data[[var_i]])) +
    # geom_boxplot(outlier.alpha=0)+
    geom_beeswarm(alpha = .2,
                  aes(color = PAD)) +
    stat_summary(
      fun.data = mean_sdl,
      fun.args = list(mult = 1)#,
      # color = "darkolivegreen1"
    ) +
    geom_signif(
      comparisons = list(c(1, 2)),
      annotations = gauss_out |>
        filter(Variable == var_i) |>
        pull(p) |>
        markSign()
    ) +
    scale_y_continuous(
      paste(var_i, "\n(Mean\u00b1SD)"),
      expand = expansion(
        mult = c(0.1, .1)
      )
    ) +
    scale_color_manual(values = pad_colors)
  print(plot_out)
  cat("<br>\n\n")
}
```

\newpage

### Ordinal Variables

```{r}
cat("<br>\n\n")
ord_out <- compare2numvars(
  rawdata,
  dep_vars = ord_vars$names,
  indep_var = "PAD",
  gaussian = FALSE,
  add_n = TRUE,
  round_desc = 3
)
ord_out |>
  select(-desc_all) |>
  flextable() |>
  set_table_properties(
    width = 1,
    layout = "autofit"
  )
cat("<br>\n\n")
```

```{r}
for (var_i in ord_vars$names) {
  plot_out <-
    ggplot(rawdata, aes(PAD, .data[[var_i]])) +
    geom_boxplot(outlier.alpha = 0) +
    geom_beeswarm(alpha = .2,
                  aes(color = PAD)) +
    geom_signif(
      comparisons = list(c(1, 2)),
      annotations = ord_out |>
        filter(Variable == var_i) |>
        pull(p) |>
        markSign()
    ) +
    scale_y_continuous(expand = expansion(
      mult = c(0.1, .1)
    )) +
    scale_color_manual(values = pad_colors)
  if (str_detect(var_i, "Tri|Lox|LP|hsC|BNP|Alb")) {
    plot_out <- plot_out +
      scale_y_log10(
        expand = expansion(
          mult = c(0.1, .1)
        ),
        breaks = logrange_15,
        minor_breaks = logrange_123456789,
        labels = prettyNum
      )
  }
  print(plot_out)
  cat("<br>\n\n")
}
```

\newpage

### Factor Variables

```{r}
cat("<br>\n\n")
fact_out <- compare2qualvars(
  rawdata,
  dep_vars = fact_vars$names,
  indep_var = "PAD", pretext = TRUE
)
fact_out |>
  select(-desc_all) |>
  flextable() |>
  padding(~ p == " ", padding.left = 20) |>
  bg(~ p != " ", bg = "grey90") |>
  set_table_properties(
    width = 1,
    layout = "autofit"
  )
cat("<br>\n\n")
```

```{r}
#| fig.height: 4
#| fig.width: 3
for (var_i in fact_vars$names) {
  plot_out1 <-
    rawdata |>
    # filter(!is.na(!!sym(var_i))) |>
    drop_na(all_of(var_i)) |>
    ggplot(aes(PAD, fill = .data[[var_i]])) +
    geom_bar(position = "fill") +
    geom_signif(
      comparisons = list(c(1, 2)),
      aes(y = 1.05),
      annotations = fact_out |>
        filter(Variable == var_i) |>
        pull(p) |>
        markSign()
    ) +
    scale_y_continuous(
      breaks = seq(0, 1, .2),
      labels = scales::percent,
      expand = expansion(
        mult = c(0.1, .1)
      )
    ) +
    theme(legend.position = "bottom") +
    scale_fill_manual(values = pad_colors)
  plot_out2 <-
    rawdata |>
    # filter(!is.na(!!sym(var_i))) |>
    drop_na(all_of(var_i)) |>
    ggplot(aes(.data[[var_i]], fill = PAD)) +
    geom_bar(position = "fill") +
    geom_signif(
      comparisons = list(c(1, 2)),
      aes(y = 1.05),
      annotations = fact_out |>
        filter(Variable == var_i) |>
        pull(p) |>
        markSign()
    ) +
    scale_y_continuous(
      breaks = seq(0, 1, .2),
      labels = scales::percent,
      expand = expansion(
        mult = c(0.1, .1)
      )
    ) +
    theme(legend.position = "bottom") +
    scale_fill_manual(values = pad_colors)
  print(plot_out1+plot_out2)
  cat("<br>")
}
```

\newpage

## Correlations

```{r}
#| fig.height: 8
if (internal_version) {
  rawdata |>
    select(all_of(gauss_vars$names)) |>
    rename_with(~ str_wrap(.x, 10)) |>
    ggpairs(
      lower = list(
        continuous =
          wrap("points", alpha = 0.1)
      ),
      upper = list(
        continuous =
          wrap("cor",
               size = 3,
               method = "pears"
          )
      )
    ) +
    theme(strip.text = element_text(
      size = 6,
      lineheight = 0.9, face = "bold"
    ))
  cat("<br>\n\n")
  rawdata |>
    select(all_of(ord_vars$names)) |>
    rename_with(~ str_wrap(.x, 10)) |>
    ggpairs(
      lower = list(
        continuous =
          wrap("points", alpha = 0.1)
      ),
      upper = list(
        continuous =
          wrap("cor",
               size = 3,
               method = "spearman"
          )
      )
    ) + theme(strip.text = element_text(
      size = 6,
      lineheight = 0.9, face = "bold"
    ))
  cat("<br>\n\n")
}
```

```{r}
# only ord_vars here for the sake of simplicity...
cortestR(
  rawdata |>
    select(all_of(ord_vars$names)),
  method = "spear"
) |>
  flextable() |>
  set_table_properties(
    width = 1,
    layout = "autofit"
  )
cat("<br>\n\n")
cortestR(
  rawdata |>
    select(all_of(ord_vars$names)),
  method = "spear"
) |>
  rownames_to_column("Reference") |>
  select(1:2) |>
  slice(-1) |> 
  flextable() |>
  set_table_properties(
    width = 1,
    layout = "autofit"
  )
cat("<br>\n\n")

```

## Prediction analysis

### All univariable models

```{r}
cat("<br>\n\n")
OR_plotdata_uni <- tibble(
  Predictor = NA_character_,
  OR_univ = NA_real_,
  CI_low_univ = NA_real_,
  CI_high_univ = NA_real_,
  p_univ = NA_real_,
  Significance_univ = NA_character_,
  .rows = 0
)

predvars <- c(
  gauss_vars$bticked,
  ord_vars$bticked,
  fact_vars$bticked[-1]
)

for (var_i in predvars) {
  glmformula <- paste("PAD~", var_i) |>
    as.formula()
  logreg_out <- glm(glmformula,
                    family = binomial(), data = rawdata
  )
  ORs <- exp(logreg_out$coefficients)
  CIs <- exp(confint(logreg_out))
  sum_out <- summary(logreg_out)
  OR_plotdata_uni <- add_row(OR_plotdata_uni,
                             Predictor = names(ORs)[-1],
                             OR_univ = ORs[-1],
                             CI_low_univ = CIs[-1, 1],
                             CI_high_univ = CIs[-1, 2],
                             p_univ = sum_out$coefficients[-1, 4]
  )
}

OR_plotdata_uni <-
  mutate(OR_plotdata_uni,
         Predictor = Predictor |>
           str_replace_all(
             c(
               "`" = "",
               "yes$" = ": yes",
               "female" = ": female" # "_"=" ",
             )
           ),
         Significance_univ = p_univ |> markSign()
  )
OR_plotdata_uni |>
  select(-p_univ) |>
  rowwise() |>
  mutate(across(2:4, ~ roundR(.x, 4))) |>
  ungroup() |>
  arrange(Predictor) |>
  flextable() |>
  bold(~ str_detect(Significance_univ, "\\*"),
       j = 1:2
  ) |>
  set_table_properties(
    width = 1,
    layout = "autofit"
  )
cat("<br>\n\n")
```

```{r}
#| fig.height: 8
baseplot <- ggplot(
  OR_plotdata_uni,
  aes(x = Predictor, y = OR_univ)
) +
  geom_pointrange(aes(
    ymin = CI_low_univ,
    ymax = CI_high_univ
  )) +
  coord_flip() +
  geom_hline(
    yintercept = 1,
    linewidth = .2, linetype = 2
  ) +
  scale_y_log10(
    breaks = logrange_15,
    minor_breaks = logrange_123456789,
    guide = guide_axis(n.dodge = 2),
    labels = prettyNum
  ) +
  geom_text(aes(label = Significance_univ),
            vjust = 1.5, color = "red"
  ) +
  ggtitle("OddsRatios shown on log-scale") +
  xlab(NULL)
baseplot
```

\newpage

### Multivariable model

```{r}
cat("<br>\n\n")
# predvars
predvars_m <- ColSeeker(rawdata,
                        namepattern = ".",
                        exclude = c(
                          "PLA", "LP", "CRP",
                          "PAD", "Heal",
                          "ID", "BMI",
                          "Siz", "Gew",
                          " .ln."
                        )
) |>
  pluck("bticked")
glmformula_m <-
  reformulate(predvars_m,response = "PAD")
pred_data <- rawdata |>
  select(all_of(bt(predvars_m, remove = TRUE)), PAD) |>
  drop_na()
logreg_out_m <- glm(glmformula_m,
                    family = binomial(), data = pred_data
)
# Anova(logreg_out_m)
ORs_m <- exp(logreg_out_m$coefficients)
CIs_m <- exp(confint(logreg_out_m))
sum_out_m <- summary(logreg_out_m)
OR_plotdata_m <- tibble(
  Predictor = names(ORs_m)[-1],
  OR = ORs_m[-1],
  CI_low = CIs_m[-1, 1],
  CI_high = CIs_m[-1, 2],
  p = sum_out_m$coefficients[-1, 4],
  Significance = markSign(p)
) |>
  arrange(OR) |>
  mutate(Predictor = fct_inorder(Predictor))

OR_plotdata_m |>
  mutate(across(2:4, ~ roundR(.x, 4)),
         p = formatP(p, mark = TRUE)
  ) |>
  rowwise() |>
  mutate(across(2:4, ~ roundR(.x, 4))) |>
  ungroup() |>
  select(-p) |>
  flextable() |>
  bold(~ str_detect(Significance, "\\*"),
       j = 1:2
  ) |>
  set_table_properties(
    width = 1,
    layout = "autofit"
  )
cat("<br>\n\n")
```

```{r}
#| fig.height: 8
baseplot_m <- ggplot(
  OR_plotdata_m,
  aes(x = Predictor, y = OR)
) +
  geom_pointrange(aes(
    ymin = CI_low,
    ymax = CI_high
  )) +
  coord_flip() +
  geom_hline(
    yintercept = 1,
    linewidth = .2, linetype = 2
  ) +
  scale_y_log10(
    breaks = logrange_15,
    minor_breaks = logrange_123456789,
    guide = guide_axis(n.dodge = 2),
    labels = prettyNum
  ) +
  geom_text(aes(label = Significance),
            vjust = 1.5, color = "red"
  ) +
  ggtitle("OddsRatios shown on log-scale") +
  xlab(NULL)
baseplot_m
```

```{r}
pred_data$pPAD <- predict(logreg_out_m,
                          type = "response"
)
# pROC
roc_out <- roc(
  response = pred_data$PAD,
  predictor = pred_data$pPAD
)
youden <- pROC::coords(roc_out, x = "best", best.method = "youden")
# youden
ggroc(roc_out, legacy.axes = TRUE) +
  geom_abline(slope = 1, intercept = 0) +
  geom_point(
    x = 1 - youden$specificity,
    y = youden$sensitivity,
    color = "red",
    size = 2
  ) +
  annotate(
    geom = "label",
    x = 1 - youden$specificity,
    y = youden$sensitivity,
    label = paste(
      "Spec.:", round(youden$specificity, 2),
      "\nSens.:", round(youden$sensitivity, 2)
    ),
    vjust = 1.1,
    hjust = 0,
    color = "red"
  )

ggplot(pred_data, aes(PAD, pPAD)) +
  geom_boxplot(outlier.alpha = 0) +
  scale_y_continuous(breaks = seq(0, 1, .1)) +
  ggbeeswarm::geom_beeswarm(alpha = .25) +
  geom_hline(
    yintercept = c(
      youden$threshold,
      .5
    ),
    color = c("red", "blue"),
    linetype = 2:3
  )
```
