---
title: "Cute penguin models"
format: 
  docx:
    reference-doc:  template_class.docx
    toc: true
  # pdf:
    # toc: true
editor: visual
execute: 
  echo: false
  warning: false
  output: asis
fig-dpi: 300
fig-width: 6
fig-height: 8
tbl-cap-location: top
---

```{r setupstuff}
pacman::p_load(conflicted,tidyverse,wrappedtools,
               rlist, flextable, ggsignif,
               patchwork, ggbeeswarm, broom, car,
               easystats, multcomp, emmeans,
               palmerpenguins
               )
conflicts_prefer(dplyr::filter, 
                 dplyr::select,
                 palmerpenguins::penguins)
set_flextable_defaults(big.mark = " ", 
                       font.size = 9, 
                       theme_fun = theme_vanilla,
                       padding.bottom = 1, 
                       padding.top = 3,
                       padding.left = 3,
                       padding.right = 4#,
                       # font.family = "Roboto Condensed"
)
theme_set(theme_bw())
# theme_update(
#   plot.title = element_text(family = "Roboto Condensed"),
#   plot.caption = element_text(family = "Roboto Condensed"),
#   axis.title = element_text(family = "Roboto Condensed"),
#   axis.text = element_text(family = "Roboto Condensed"),
#   strip.text = element_text(family = "Roboto Condensed"))
gmx_colors <- c("#AE177C", "#20346E", "#784CD8",
                "#E70B76", "#FDCCF8") 
```

```{r dataprep}
rawdata <- penguins |>
  drop_na() |> 
  mutate(year = factor(year))
Adelies <- rawdata |> 
  filter(species == "Adelie") |> 
  mutate(`body weight [kg]` = body_mass_g/1000)
```

# Regression-like:

## Model flipper length depending on weight

to dos:

- vizualize the data for the task (points, smoothies)

- build a model and put into variable

- print model output (tidy/ flextable?)

- test significance of model terms

- print significance

- check model??

- if suitable, visualize results

```{r}
rawdata |> 
  ggplot(aes(body_mass_g)) +
  geom_density()
cat("<br>\n\n")
Adelies |> 
  ggplot(aes(body_mass_g)) +
  geom_density()
cat("<br>\n\n")
rawdata |> 
  ggplot(aes(body_mass_g,flipper_length_mm)) +
  geom_point() +
  geom_smooth(color = "darkgreen") +
  geom_smooth(method = "lm")
cat("<br>\n\n")
Adelies |> 
  ggplot(aes(body_mass_g,flipper_length_mm)) +
  geom_point() +
  geom_smooth(color = "darkgreen") +
  geom_smooth(method = "lm")
cat("<br>\n\n")
flipper_by_weight <- 
  lm(flipper_length_mm~body_mass_g,
     data = rawdata)
tidy(flipper_by_weight) |> 
  select(1:2) |> 
  flextable() |> 
  set_table_properties(width = .6,layout = 'autofit')
cat("<br>\n\n")
anova(flipper_by_weight) |> 
  tidy() |> 
  mutate(p.value = formatP(p.value,ndigits = 5,
                           mark = TRUE)) |> 
  flextable() |> 
  set_table_properties(width = .8,layout = 'autofit')
cat("<br>\n\n")


```

## prediction

```{r}
#create NAs
set.seed(2026)
NA_positions <- 
  sample(1:nrow(rawdata), 20,replace = FALSE) |> 
  sort()
penguins_with_NAs <- rawdata
penguins_with_NAs[NA_positions, "flipper_length_mm"] <- NA
# penguins_with_NAs[NA_positions, "flipper_length_mm"] |> 
#   flextable()
# cat("<break>\n\n")

# model
flipper_model <- 
  lm(flipper_length_mm~body_mass_g,
     data = penguins_with_NAs)
 penguins_with_NAs <- 
   mutate(penguins_with_NAs,
          flipper_length_mm = 
            case_when(is.na(flipper_length_mm) ~
                        flipper_model$coefficients[1] +
                        flipper_model$coefficients[2] * 
                          body_mass_g,
                      .default = flipper_length_mm 
                      )
          )
penguins_with_NAs[NA_positions, "flipper_length_mm"]
real_vs_predicted <- 
  tibble(real = rawdata$flipper_length_mm[NA_positions],
         predicted = penguins_with_NAs$flipper_length_mm[NA_positions]
  )
ggplot(real_vs_predicted, 
       aes(x = real, y = predicted)) +
  geom_point()
```

## Change in units -\> change in slope

```{r}
rawdata <-    
   mutate(rawdata,
          `body weight [kg]` = body_mass_g/1000)
 

lm_kg <- 
  lm(flipper_length_mm ~ `body weight [kg]`,
     data = rawdata)
lm_kg_a <- 
  lm(flipper_length_mm ~ `body weight [kg]`,
     data = Adelies)
anova_kg <- anova(lm_kg) |> 
  tidy()
lm_kg  |> 
  tidy() |> 
  mutate(p.value = formatP(p.value,ndigits = 5,
                           mark = TRUE)) |> 
  flextable() |> 
  set_table_properties(width = 1,layout = 'autofit')
cat('<br>\n\n')
if (interactive()) {
  x11()
}
check_model(lm_kg)
cat("<br>\n\n")
check_model(lm_kg_a)
if (interactive()) {
  dev.off()
}

rawdata |> 
  ggplot(aes(`body weight [kg]`,flipper_length_mm)) +
  geom_point() +
  geom_smooth(method = "lm",
              color = gmx_colors[1], fill = gmx_colors[2]) +
  annotate(geom = "label",
           x = 5.75, y = 180,
           size = 3,
           label = paste('Increase flipper length\nper kg body weight (Slope):\n', 
                          roundR(lm_kg$coefficients[2],3),
                          " p",formatP(anova_kg$p.value[1],
                                      pretext = T,mark = T)))
cat("<br>\n\n")

```

## Model bill depth depending on weight

```{r}
rawdata |> 
  ggplot(aes(body_mass_g, bill_depth_mm)) +
  geom_point() +
  geom_smooth(color = "darkgreen") +
  geom_smooth(method = "lm")
cat("<br>\n\n")
bill_by_weight <- 
  lm(bill_depth_mm ~ body_mass_g,
     data = rawdata)
tidy(bill_by_weight) |> 
  select(1:2) |> 
  flextable() |> 
  set_table_properties(width = .6,layout = 'autofit')
cat("<br>\n\n")
anova(bill_by_weight) |> 
  tidy() |> 
  mutate(p.value = formatP(p.value,ndigits = 5,
                           mark = TRUE)) |> 
  flextable() |> 
  set_table_properties(width = .6,layout = 'autofit')
cat("<br>\n\n")
if (interactive()) {
  x11()
}
check_model(bill_by_weight)
if (interactive()) {
  dev.off()
}
```

Adjust for species and sex first?

```{r}
rawdata |> 
  ggplot(aes(body_mass_g,bill_depth_mm)) +
  geom_point(aes(color = species, shape = sex)) +
  geom_smooth(color = "darkgreen") +
  geom_smooth(method = "lm")
cat("<br>\n\n")
bill_by_sp_sex_weight <- 
  lm(bill_depth_mm~species +sex +body_mass_g,
     data = rawdata)
tidy(bill_by_sp_sex_weight) |> 
  select(1:2) |> 
  flextable() |> 
  set_table_properties(width = .6,layout = 'autofit')
cat("<br>\n\n")
anova(bill_by_sp_sex_weight) |> 
  tidy() |> 
  mutate(p.value = formatP(p.value,ndigits = 5,
                           mark = TRUE)) |> 
  flextable() |> 
  set_table_properties(width = .6,layout = 'autofit')
cat("<br>\n\n")

Anova(bill_by_sp_sex_weight) |> 
  tidy() |> 
  mutate(p.value = formatP(p.value,ndigits = 5,
                           mark = TRUE)) |> 
  flextable() |> 
  set_table_properties(width = .6,layout = 'autofit')
cat("<br>\n\n")

check_model(bill_by_sp_sex_weight)

rawdata |> 
  ggplot(aes(body_mass_g,bill_depth_mm)) +
  geom_point(aes(color = species, shape = sex)) +
  geom_abline(
    intercept = bill_by_sp_sex_weight$coefficients[1],
    slope = bill_by_sp_sex_weight$coefficients["body_mass"],
    color = "darkorange3",
    linewidth = 2) +
  geom_smooth(method = "lm")
cat("<br>\n\n")


```

# Anova-like:

## flipper length by species

```{r}
rawdata |> 
  ggplot(aes(species,flipper_length_mm)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_beeswarm(alpha = .2)
flipper_by_species <- 
  lm(flipper_length_mm ~ species,
     data = rawdata)
tidy(flipper_by_species) |> 
  select(1:2, p.value)|> 
  mutate(p.value = formatP(p.value,ndigits = 3, mark = T)) |> 
  flextable() |> 
  set_table_properties(width = 1,layout = 'autofit')
cat("<br>\n\n")
anova(flipper_by_species) |> 
  tidy() |> 
  mutate(p.value = formatP(p.value,4)) |> 
  flextable() |> 
  set_table_properties(width = 1,layout = 'autofit')
cat("<br>\n\n")
if (interactive()) {
  x11()
}
check_model(flipper_by_species)
if (interactive()) {
  dev.off()
}
posthoc <- 
  pairwise.t.test(rawdata$flipper_length_mm,rawdata$species,
                p.adjust.method = "fdr") |> 
  tidy()

posthoc |> 
  # as_tibble(rownames = "Reference")|> 
  mutate(p.value = formatP(p.value,ndigits = 5,
                           mark = TRUE)) |>
  flextable() |>
  set_table_properties(width = 1,layout = 'autofit')

posthoc2 <- 
  summary(glht(model = flipper_by_species,
             linfct = mcp(species = 'Tukey')))

posthoc2$test$pvalues

rawdata |> 
  ggplot(aes(species,flipper_length_mm)) +
#  geom_boxplot(outlier.alpha = 0) +
  geom_beeswarm(alpha = .2) +
  stat_summary(color = "darkblue",
               fun.data = mean_cl_normal) +
  geom_signif(comparisons = list(c(1,2),c(1,3), c(2,3)),
              annotations = #markSign(posthoc2$test$pvalues),
              paste("p",posthoc2$test$pvalues |>
              formatP(3,pretext = T, mark= T)),
              y_position = c(240,250,240),
              extend_line = -.005) +
  scale_y_continuous(
    name = "flipper length\nmean \u00b1 95% CI",
    expand = expansion(mult = .1)) 
```

## flipper by species and sex (and interaction?)

```{r}
rawdata |> 
  ggplot(aes(species,flipper_length_mm, color = sex)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_beeswarm(alpha = .25,
                dodge.width = .75) +
  scale_color_manual(values = gmx_colors)

rawdata |> 
  ggplot(aes(species,flipper_length_mm, color = sex)) +
  geom_beeswarm(alpha = .25,
                dodge.width = .25) +
  stat_summary(position = position_dodge(width = .25),
               fun.data = mean_cl_normal) +
  stat_summary(fun = mean,geom = "line", 
               aes(group = sex),
               position = position_dodge(width = .25)) +
  ylab("Flipper length (mm) \u00b1 95% CI") +
  scale_color_manual(values = gmx_colors)


flipper_by_species_star_sex <- 
  lm(flipper_length_mm ~ species*sex,
     data = rawdata)
if (interactive()) {
  x11()
}
check_model(flipper_by_species_star_sex)
rawdata |> 
  slice(c(124,272,313)) |> 
  select(species, sex, flipper_length_mm,body_mass_g)
tidy(flipper_by_species_star_sex) |> 
  select(1:2) |> 
  flextable() |> 
  set_table_properties(width = .7,layout = 'autofit')
cat("<br>\n\n")
Anova_out <- 
  Anova(flipper_by_species_star_sex,type = 3) |> 
  tidy() |> 
  mutate(p.value = formatP(p.value,4))  
Anova_out |> 
  flextable() |> 
  set_table_properties(width = 1,layout = 'autofit')
cat("<br>\n\n")
posthoc <- summary(glht(model = flipper_by_species_star_sex,
             linfct = mcp(species = 'Tukey',
                        sex = "Tukey"))) |> 
  tidy()
posthoc |> 
  mutate(adj.p.value = formatP(adj.p.value, 5, mark = T)) |> 
  flextable()
tempdata <- mutate(
  rawdata,
  species_sex = interaction(species, sex)
)
flipper_by_species_star_sex2 <- 
  lm(flipper_length_mm ~ species_sex,
     data = tempdata)
tidy(flipper_by_species_star_sex2) |> 
  select(1:2) |> 
  flextable() |> 
  set_table_properties(width = .7,layout = 'autofit')
cat("<br>\n\n")
Anova_out2 <- 
  Anova(flipper_by_species_star_sex2,type = 3) |> 
  tidy() |> 
  mutate(p.value = formatP(p.value,4))  
Anova_out2 |> 
  flextable() |> 
  set_table_properties(width = 1,layout = 'autofit')
cat("<br>\n\n")

posthoc_star2 <- summary(glht(model = flipper_by_species_star_sex2,
             linfct = mcp(species_sex = 'Tukey'))) |> 
  tidy()

posthoc_star2 |> 
  mutate(p.value = formatP(adj.p.value),
         estimate = roundR(estimate,3)) |> 
  select(contrast,estimate,p.value) |> 
  flextable()

pairwise.t.test(tempdata$flipper_length_mm,
                tempdata$species_sex,
                p.adjust.method = "n")

nested_p <- emmeans(flipper_by_species_star_sex,
                    ~ sex | species) |> 
  pairs() |> tidy()
nested_p2 <- emmeans(flipper_by_species_star_sex,
                    ~ species | sex) |> 
  pairs() |> tidy()
rawdata |> 
  ggplot(aes(species,flipper_length_mm, color = sex)) +
  geom_beeswarm(alpha = .25,
                dodge.width = .25) +
  stat_summary(position = position_dodge(width = .25),
               fun.data = mean_cl_normal) +
  stat_summary(fun = mean,geom = "line", 
               aes(group = sex),
               position = position_dodge(width = .25)) +
  ylab("Flipper length (mm)\nmean\u00b1 95% CI") +
  geom_signif(comparisons = list(c(1,2),c(1,3), c(2,3)),
              annotations = posthoc$adj.p.value[-4] |> 
                formatP(),#markSign(),
              y_position = c(240,250,240),
              extend_line = -.005,
              color = "black") +
  scale_y_continuous(expand = expansion(mult = .1)) +
  scale_color_manual(
    name = paste('sex\n p ', 
                 Anova_out |> 
                   filter(term == "sex") |> 
                   pull(p.value) |> 
                   formatP(pretext = TRUE)),
    values = gmx_colors) +
  annotate(geom = "label", x = 3.75, y = 215,
           label = paste("interaction\nsex*species\np",
                         Anova_out |> 
                           filter(term == "species:sex") |> 
                           pull(p.value) |> 
                           formatP(pretext = TRUE))) +
  scale_x_discrete(expand = expansion(add = c(.5,1.25))) +
  geom_signif(xmin = c(.9,1.9, 2.9), 
              xmax = c(1.1, 2.1,3.1), 
              y_position = c(220,225, 235), 
              annotations = paste("p",
                                  nested_p$p.value |> 
                                    formatP(pretext = T, 
                                            mark = T)),
              color = "black", textsize = 2.5)

flipper_by_species_plus_sex <- 
  lm(flipper_length_mm ~ species +sex,
     data = rawdata)
tidy(flipper_by_species_plus_sex) |> 
  select(1:2) |> 
  flextable() |> 
  set_table_properties(width = 1,layout = 'autofit')
cat("<br>\n\n")

anova(flipper_by_species_plus_sex, 
      flipper_by_species_star_sex) |> 
  tidy() |> 
  flextable()


Anova(flipper_by_species_plus_sex,type = 2) |> 
  tidy() |> 
  mutate(p.value = formatP(p.value,4)) |> 
  flextable() |> 
  set_table_properties(width = 1,layout = 'autofit')
cat("<br>\n\n")

```

# combined:

## flipper len by species/sex/weight

```{r}
rawdata |>   
  ggplot(aes(body_mass_g,flipper_length_mm)) +
  geom_point(alpha = .2) +
  # geom_smooth(color = "darkgreen") +
  geom_smooth(method = "lm") +
  facet_grid(cols = vars(sex),
             rows = vars(species),
             margins = TRUE)
cat("<br>\n\n")
flipper_with_stars <- 
  lm(flipper_length_mm~(body_mass_g +species +sex)^2,  # this calls for all  2-way interactions
     data = rawdata)
tidy(flipper_with_stars) |> 
  select(1:2) |> 
  flextable() |> 
  set_table_properties(width = 1,layout = 'autofit')
cat("<br>\n\n")
Anova(flipper_with_stars,type = 3) |> 
  tidy() |> 
  mutate(p.value = formatP(p.value,4,mark = T)) |> 
  flextable() |> 
  set_table_properties(width = 1,layout = 'autofit')
cat("<br>\n\n")

flipper_with_plus <- 
  lm(flipper_length_mm~body_mass_g +species +sex,
     data = rawdata)
tidy(flipper_with_plus) |> 
  select(1:2) |> 
  flextable() |> 
  set_table_properties(width = 1,layout = 'autofit')
cat("<br>\n\n")
Anova(flipper_with_plus,type = 2) |> 
  tidy() |> 
  mutate(p.value = formatP(p.value,4)) |> 
  flextable() |> 
  set_table_properties(width = 1,layout = 'autofit')
cat("<br>\n\n")
anova(flipper_with_stars, flipper_with_plus) |> 
  tidy() |> 
  flextable() |> 
  set_table_properties(width = 1,layout = 'autofit')
cat('<br>\n\n')


flipper_with_sex_star_species <- 
  lm(flipper_length_mm~body_mass_g + species * sex,
     data = rawdata)
tidy(flipper_with_sex_star_species) |> 
  select(1:2) |> 
  flextable() |> 
  set_table_properties(width = 1,layout = 'autofit')
cat("<br>\n\n")
Anova(flipper_with_sex_star_species,type = 3) |> 
  tidy() |> 
  mutate(p.value = formatP(p.value,4)) |> 
  flextable() |> 
  set_table_properties(width = 1,layout = 'autofit')
cat("<br>\n\n")
anova(flipper_with_stars, flipper_with_sex_star_species) |> 
  tidy() |> 
  flextable() |> 
  set_table_properties(width = 1,layout = 'autofit')
cat('<br>\n\n')
if (interactive()) {
  x11()
}
check_model(flipper_by_species_star_sex)
if (interactive()) {
  dev.off()
}
```

create formulas programmatically

```{r}
pred_vars <- ColSeeker(namepattern = c("^s[ep]","mass"))
star_formula <- paste("flipper_len",
                      paste(pred_vars$names, collapse = "*"),
                      sep = "~")
star_formula
plus_formula <- paste("flipper_len",
                      paste(pred_vars$names, collapse = "+"),
                      sep = "~")
plus_formula
```
