---
title: "20260901_exercise_015"
author: "Sven Meinerz"
format: 
  docx:
    reference-doc:  template_class.docx
    toc: true
    toc-depth: 3
    toc-title: Table of Content
editor: visual
execute: 
  echo: false # Does not show code in rendered document
  warning: false # Common option, hides warnings; errors will still break the code
  output: asis # as is, rendering will not round numbers etc. Leave the output as it is
fig-dpi: 300
fig-width: 6
fig-height: 8
---

## Setup

Load in all needed packages and solve conflicts

```{r}
pacman::p_load(
 car, conflicted, tidyverse, ggplot2, wrappedtools, flextable,
 broom, DescTools
)
conflicts_prefer(
  dplyr::select,
  dplyr::filter,
  ggplot2::mean_cl_boot,
  modelbased::standardize
)
```

## Read in the dataset

```{r}
rawdata <- as_tibble(carData::Greene) |> drop_na() |> 
  mutate(rawdata,
    nation = fct_lump_min(nation, min = 15) |>
      fct_infreq(),
    location = fct_recode(location, "Other" = "other") |>
      fct_infreq(),
    judge = fct_infreq(judge),
    language = fct_infreq(language)
  )
head(rawdata)
```

The dataset seems fine, no need for data clean-up or changing of datatypes.

## Data exploration

```{r}
#| echo: false
ggplot(rawdata, aes(nation, fill = fct_rev(decision))) +
  geom_bar(position = "fill") +
  scale_y_continuous("Frequency", labels = scales::percent) +
  scale_fill_manual("decision",
                    values = c("lightblue", "salmon"),
                    labels = c("yes", "no")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
#| echo: false
ggplot(rawdata, aes(language, fill = fct_rev(decision))) +
  geom_bar(position = "fill") +
  scale_y_continuous("Frequency", labels = scales::percent) +
  scale_fill_manual("decision",
                    values = c("lightblue", "salmon"),
                    labels = c("yes", "no")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
#| echo: false
ggplot(rawdata, aes(judge, fill = decision)) +
  geom_bar(position = "fill") +
  scale_y_continuous("Frequency", labels = scales::percent) +
  scale_fill_manual("decision",
                    values = c("lightblue", "salmon"),
                    labels = c("yes", "no")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
#| echo: false
ggplot(rawdata, aes(location, fill = decision)) +
  geom_bar(position = "fill") +
  scale_y_continuous("Frequency", labels = scales::percent) +
  scale_fill_manual("decision",
                    values = c("lightblue", "salmon"),
                    labels = c("yes", "no")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

As some countries only have a few data points, buiulding a model on those data would be overfitting. To amend this situation we can group all countries with less than 10 caeses inside a "catch all" category.

```{r}
#| echo: true
rawdata <- mutate(rawdata, nation = fct_lump_min(nation, min = 10))
table(rawdata$nation)
```

## Modelling

Defining all predictor variables

```{r}
predvar <- ColSeeker(data = rawdata,
                     namepattern = c("jud", "atio", "lan"))
```

### Univariable Model

First lets try building a univariable model.

```{r}
report <- tibble(
  predictor = character(),
  statistic = numeric(),
  df = numeric(),
  p.value = character()
)

for (pred_i in seq(predvar$names)) {
  
  # Build the model
  form = reformulate(predvar$names[pred_i], response = "decision")
  log_reg_out <- glm(
    form, 
    family = binomial(),
    data = rawdata
  )
  
  log_reg_out |> tidy() |> 
    mutate(p.value = formatP(p.value, ndigits = 5, mark = TRUE)) |>
    flextable() |> set_table_properties(layout = "autofit") |> 
    flex2rmd()
  sum_out <- tidy(log_reg_out)
  
  Anova_out_uni <- Anova(log_reg_out, type = 2) |> tidy() |>
  mutate(p.value = formatP(p.value, ndigits = 5, mark = TRUE))
  
  report <- report |> add_row(predictor = predvar$names[pred_i],
                  statistic = Anova_out_uni$statistic, 
                  df = Anova_out_uni$df, 
                  p.value = Anova_out_uni$p.value)
  
  ORs <- exp(log_reg_out$coefficients)
  CIs <- exp(confint(log_reg_out))
  
  # Structure for visualization
  OR_plotdata <- tibble(
  Predictor = names(ORs)[-1],
  OR = ORs[-1],
  CI_low = CIs[-1, 1],
  CI_high = CIs [-1, 2],
  p = sum_out$p.value[-1],
  Significance = markSign(p),
  label = paste(Predictor, Significance) |> fct_inorder()
  )
  # Plotting
  temp_plot <- ggplot(OR_plotdata, aes(x = Predictor, y = OR)) +
  geom_pointrange(aes(ymin = CI_low, ymax = CI_high)) +
  geom_hline(yintercept = 1, linewidth = .5, linetype = 2) +
  coord_flip() +
  scale_y_log10(
    breaks = logrange_1,
    minor_breaks = logrange_123456789,
    limits = c(0.1, 2*10^4),
    labels = prettyNum) +
  geom_text(aes(label = Significance), vjust = 1.25, color = "red") +
  xlab(NULL)
  
  print(temp_plot)
}
report |> flextable() |> set_table_properties(layout = "autofit")
```

### Multivariable Model

```{r}
glm_formula <- reformulate(predvar$names, response = "decision")
log_reg_out <- glm(
  glm_formula, family = binomial(), data = rawdata
)
log_reg_out

ORs <- exp(log_reg_out$coefficients)
CIs <- exp(confint(log_reg_out))

Anova_out_multi <- Anova(log_reg_out, type = 2) |>
  tidy() |> mutate(p.value = formatP(p.value, ndigits = 5, 
                                     mark = TRUE))
Anova_out_multi |> mutate() |>
  flextable() |> 
  set_table_properties(layout = "autofit")

sum_out <- tidy(log_reg_out)

tibble(
  Formula = deparse1(glm_formula),
  AIC = AIC(log_reg_out),
  BIC = BIC(log_reg_out)) |>
  flextable() |> set_table_properties(layout = "autofit")
```

#### Visualization

```{r}
OR_plotdata <- tibble(
  Predictor = names(ORs)[-1] |>
      str_replace("([:lower:])([:upper:])", "\\1: \\2"),
  OR = ORs[-1],
  CI_low = CIs[-1, 1],
  CI_high = CIs[-1, 2],
  p = sum_out$p.value[-1],
  Significance = markSign(p),
  Label = paste(Predictor, Significance) |> 
    fct_inorder()
)

ggplot(OR_plotdata, aes(x = Predictor, y = OR)) +
  geom_pointrange(aes(ymin = CI_low, ymax = CI_high)) +
  geom_hline(yintercept = 1, linewidth = .5, linetype = 2) +
  coord_flip() +
  scale_y_log10(
    breaks = logrange_12357,
    minor_breaks = logrange_123456789,
    guide = guide_axis(n.dodge = 2),
    labels = prettyNum,
    expand = expansion(mult = .1)) +
  geom_text(aes(label = Significance), 
            vjust = 1.5, color = "red", size = 4) +
  ggtitle("OddsRatios for leave to appeal granted\nshown on log-scale") +
  labs(caption = "Results from multivariable glms,\nOddsRatios shown on log-scale") +
  xlab(NULL)
```
