---
title: "Cervical data"
format: html
editor: visual
---

# Setup

```{r}
pacman::p_load(
  wrappedtools, tidyverse, here, conflicted, ggforce, ggbeeswarm, 
  flextable,caret, tidymodels, gmodels, easystats, yardstick, kknn, 
  rpart, rpart.plot, randomForest, future.mirai
)

conflicts_prefer(
  dplyr::select,
  dplyr::filter,
)
```

# Data loading

```{r}
cervical <- readRDS("data/cervical.RDS")
```

# Normality

```{r}
normality <- cervical |> 
  summarise(
    across(.cols = where(is.numeric),
           .fns = list(
             pKS = ~ksnormal(.x,lillie = FALSE)
           )
    )
  ) |>
  pivot_longer(cols = everything(),
               names_to = "miRNA",
               values_to = "p-value") 


normality <- normality |>
  mutate(
    normal = case_when(normality$`p-value` <= 0.05 ~ 0,
                       normality$`p-value` > 0.05 ~ 1),
    normal = factor(normal, levels = c(0, 1), labels = c("no", "yes"))
  )

normality |> ggplot(aes(normal, fill = normal)) +
  geom_bar()

normality |> ggplot(aes(`p-value`)) +
  geom_histogram(binwidth = 0.01)

rm(normality)
```

# Data Splitting

```{r}
set.seed(20260909)
data_split <- cervical |>
  select(Tissuetype, contains("_")) |>
  initial_split(
  cervical,
  prop = 0.7,
  strata = Tissuetype
)
traindata <- training(data_split)
testdata <- testing(data_split)
```

# Preprocessing

NZV to remove near zero variance data, most of the time those RNA with near zero expression. Yeo-Johnson is used to un-skew the data. Highly correlated data is filtered out to reduce dimension. Spearmann is used as data is not normal distributed. Lastly, data is normalized using quantile normalization. The argument "outside" is used to prepare for the case, that test data may have data points, which are outside the range of the training data set.

```{r}
rec_rna <- recipe(
  x = traindata) |>
  update_role(Tissuetype, new_role = "outcome") |>
  update_role(contains("_"), new_role = "predictor") |>
  step_nzv(all_predictors()) |>
  step_YeoJohnson(all_predictors()) |>
  step_corr(all_predictors(), method = "spearman") |>
  step_percentile(all_predictors(), outside = "both")
```

Fit recipe and apply to data.

```{r}
traindata_s <- rec_rna |>
  prep() |>
  bake(new_data = traindata) |>
  relocate(Tissuetype, .before = 1)
testdata_s <- rec_rna |>
  prep() |>
  bake(new_data = testdata) |>
  relocate(Tissuetype, .before = 1)

# Create vector of variables which are used inside of the model
predvars <- ColSeeker(traindata_s, "\\d") 
```

# Exploratory

Visualize the effect of the training on the data.

```{r}
ks_out <- traindata_s |>
  select(-Tissuetype) |>
  summarise(across(
    everything(),
    ~ ksnormal(.x)
  )) |>
  pivot_longer(everything(), values_to = "pKS") |>
  arrange(pKS)
ks_out |>
  ggplot(aes(pKS)) +
  scale_x_continuous(
    breaks = seq(0, 1, .1),
    minor_breaks = seq(0, 1, .01)
  ) +
  geom_histogram(binwidth = .01) # bins=20)

ks_out |>
  group_by(pKS < .05) |>
  count()
traindata_s |>
  select(Tissuetype, all_of(sample(predvars$names, 16))) |>
  pivot_longer(-Tissuetype) |>
  ggplot(aes(value, fill = Tissuetype)) +
  geom_density(alpha = .2) +
  facet_wrap(facets = vars(name), scales = "free")

explore_s <- traindata_s |>
  select(-Tissuetype) |>
  summarise(across(
    everything(),
    list(
      Mean = ~ mean(.x, na.rm = T),
      Median = ~ median(.x, na.rm = T),
      SD = ~ sd(.x, na.rm = T)
    )
  )) |>
  pivot_longer(everything(),
    names_to = c("miRNA", ".value"),
    names_pattern = "(.+)_([MS].+)"
  )

ggplot(
  traindata_s,
  aes(.data[["let_7a"]],
    .data[["let_7c"]],
    color = Tissuetype
  )
) +
  geom_point(alpha = .5) +
  geom_mark_hull()
```

# k nearest neighbor

## Train Model

Preprocessing can also be done inside the workflow, which might be more convenient.

```{r}
knn_spec <- nearest_neighbor(neighbors = 3) |>
  set_engine("kknn") |>
  set_mode("classification")

# First workflow can be applied to scaled, preprocessed data
knn_workflow <- workflow() |>
  add_model(knn_spec) |>
  add_formula(Tissuetype ~ .)

# Used on unbaked data. Often used on more complicated data. This would be more automated.
knn_workflow2 <- workflow() |>
  add_model(knn_spec) |>
  add_recipe(rec_rna)
```

## Fit the model to the training data.

```{r}
knn_fit <- knn_workflow |>
  fit(data = traindata_s)

knn_fit2 <- knn_workflow2 |>
  fit(data = traindata)
```

## Predict Classes

```{r}
predictions <- knn_fit |>
  predict(new_data = testdata_s)

predictions2 <- knn_fit2 |>
  predict(new_data = testdata)

prob_preds <- knn_fit |>
    predict(new_data = testdata_s, type = "prob")

# Combine with original data
results <- testdata_s |>
  select(Tissuetype) |>
  bind_cols(predictions, prob_preds)
```

## Evaluate model by table/figures

```{r}
results |>
  pivot_longer(
    cols = where(is.numeric),
    names_to = "prob_class",
    values_to = "prob"
  ) |>
  ggplot(aes(prob_class, prob)) +
  geom_violin() +
  geom_beeswarm(cex = .5, alpha = .25) +
  facet_grid(
    rows = vars(Tissuetype),
    labeller = "label_both"
  )

CrossTable(results$.pred_class, results$Tissuetype,
           prop.chisq = FALSE, prop.t = FALSE,
           format = "SPSS", fisher = FALSE)

caret::confusionMatrix(results$.pred_class,
                       reference = results$Tissuetype,
                       positive = "Tumor")

yardstick::accuracy(
  data = results,
  truth = Tissuetype,
  estimate = .pred_class,
  event_label = "second"
)
yardstick::sensitivity(
  data = results,
  truth = Tissuetype,
  estimate = .pred_class,
  event_level = "second"
)
yardstick::specificity(
  data = results,
  truth = Tissuetype,
  estimate = .pred_class,
  event_level = "second"
)
yardstick::ppv(
  data = results,
  truth = Tissuetype,
  estimate = .pred_class,
  event_level = "second"
)
yardstick::npv(
  data = results,
  truth = Tissuetype,
  estimate = .pred_class,
  event_level = "second"
)
```

# Regression Tree

## Formula

```{r}
predvars <- ColSeeker(cervical, "_")
rpart_formula <- paste("Tissuetype",
  paste(predvars$names,
    collapse = "+"
  ),
  sep = "~"
)

rpart_out <- rpart(
  formula = as.formula(rpart_formula),
  data = cervical
)

prp(rpart_out,
  type = 4,
  extra = 104,
  digits = 4,
  fallen.leaves = TRUE
)

importance <- tibble(
  Variable = names(rpart_out$variable.importance),
  Importancescore = rpart_out$variable.importance
)

ggplot(importance, aes(x = Variable, y = Importancescore)) +
  geom_col() +
  coord_flip()
```

## Graphical exploration

```{r}
predvars <- ColSeeker(data = cervical, namepattern = "_")
rawplot <-
  ggplot(
    cervical,
    aes(.data[[predvars$names[1]]],
      .data[[predvars$names[2]]],
      color = Tissuetype
    )
  ) +
  geom_point()
rawplot +
  # geom_hline(yintercept = 206) +
  # geom_vline(xintercept = 500) +
  scale_y_log10() +
  scale_x_log10()
```

## Modelling

### Tidy model

```{r}
set.seed(2026)
tree_folds <- vfold_cv(
  cervical, 
  v = 5, 
  strata = Tissuetype, 
  repeats = 20)
```

one-hot encoding

```{r}
tree_recipe <- recipe(cervical) |>
  update_role(Tissuetype, new_role = "outcome") |>
  update_role(all_of(predvars$names), new_role = "predictor") #|>
  # step_dummy(all_nominal_predictors(), one_hot = TRUE)
```

tuning parameters

```{r}
tree_grid <- expand.grid(
  tree_depth = seq(2, 6, by = 2), # How many splits are allowed in model
  cost_complexity = c(0.01, 0.001), # When to stop splitting
  min_n = seq(5, 20, by = 5) # Min number for splitting, don't go too small
) |>
  as_tibble()

tree_spec <- decision_tree(
  cost_complexity = tune(),
  tree_depth = tune(),
  min_n = tune()
) |>
  set_engine("rpart", model = TRUE) |>
  set_mode("classification")

tree_workflow <- workflow() |>
  add_recipe(tree_recipe) |>
  add_model(tree_spec)
```

Initiate parallel computing for model fitting

```{r}
plan(mirai_multisession, workers = availableCores() - 2)
# Mark line 4 to 11 and execute
tictoc::tic("tuning timer")
tree_results <- tree_workflow |>
  tune_grid(
    resamples = tree_folds,
    grid = tree_grid,
    metrics = metric_set(roc_auc, accuracy, kap)
  )
tictoc::toc()
# 18.73 sec with settings from statbook
# 20 repeats : 72.78 sec
# 100 repeats: 363.01 sec (Just using higher repeats does not make the model much better)
plan(sequential)

test <- tibble(
  repeats = c(5, 20, 100),
  time = c(18.73, 75.78, 363.01)
)
test |> ggplot(aes(repeats, time)) +
  geom_point() +
  geom_smooth(method = "lm")
```

Inspect top models.
```{r}
tree_results |>
  show_best(metric = "accuracy", n = 10) |>
  select(-.config, -.estimator) |>
  mutate(across(where(is.numeric), \(x) roundR(x, 5, drop0 = TRUE))) |>
  flextable()

# Select the best parameters based on your preferred metric
tree_best <- tree_results |>
  select_best(metric = "accuracy")
```

## Visualize the tuning results
```{r}
tree_results |>
  collect_metrics() |>
  ggplot(aes(x = tree_depth, y = mean, color = as.factor(cost_complexity))) +
  geom_line(position = position_dodge(.25)) +
  geom_point(position = position_dodge(.25)) +
  scale_color_discrete("cost complexity") +
  facet_grid(
    cols = vars(.metric), rows = vars(min_n),
    labeller = "label_both"
  ) +
  theme_bw()
tree_results |>
  collect_metrics(summarize = FALSE) |>
  filter(.metric == "accuracy") |>
  ggplot(aes(x = factor(tree_depth), y = .estimate)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_beeswarm(
    alpha = 0.3,
    aes(color = .config == tree_best$.config)
  ) +
  facet_grid(
    rows = vars(min_n), cols = vars(cost_complexity), # scales = "free",
    labeller = "label_both"
  ) +
  scale_color_manual(
    "best model",
    values = c("darkgrey", "orange"),
    labels = c("no", "yes")
  ) +
  xlab("tree depth") +
  theme_bw()

```

## Updating Workflow the best model
```{r}
final_workflow <- tree_workflow |>
  finalize_workflow(tree_best)

# Fit to the entire dataset (or use test split)
final_fit <- final_workflow |>
  fit(data = cervical)

test_predicted <- final_fit |>
  augment(new_data = cervical) |>
  rename_with(\(x) str_remove(x, "^.pred_"), starts_with(".pred_"))
```

```{r}
final_out = extract_fit_engine(final_fit)
prp(final_out,
    type = 4,
    extra = 104,
    digits = 4,
    fallen.leaves = TRUE
)

importance <- tibble(
  Variable = names(final_out$variable.importance),
  Importancescore = final_out$variable.importance
) |> 
  mutate(Variable = fct_reorder(Variable, Importancescore))

ggplot(importance, aes(x = Variable, y = Importancescore)) +
  geom_col() +
  coord_flip() +
  geom_hline(yintercept = importance$Importancescore[[1]] / 2,
             color = "red",
             linetype = 2) +
  geom_hline(yintercept = importance$Importancescore[[1]] * 0.75,
             color = "blue",
             linetype = 2)
```

##Model Evaluation
```{r}
CrossTable(test_predicted$class,
  test_predicted$Tissuetype,
  prop.chisq = F, prop.t = F,
  format = "SPSS"
)

confusionMatrix(
  test_predicted$class,
  test_predicted$Tissuetype,
  positive = "Tumor"
)

tidy_metrics <- metric_set(
  accuracy, # Proportion of data that is predicted correctly
  kap, # Like accuracy but normalized by the accuracy expected by chance alone
  f_meas, # Weighted harmonic mean of precision and recall
  roc_auc, # Area under the ROC curve
  mn_log_loss # Log loss; penalizes confident predictions that are wrong
  )
tidy_metrics(test_predicted,
  truth = Tissuetype, estimate = class,
  Tumor
) |>
  flextable()
```