---
title: "HDLox"
format: html
editor: visual
---

# Tasks

- Import file use only columns up to Smoking (AB)

- use only cases with valid HDLox

- find and recode columns with 1/0 meaning yes/no

- test continuous measures for Normality / logNormality

- outcome of interest Artery Disease (AD): Peripheral/Carotid artery disease (pAVK/cAVK) (0=nein, 1=ja)

- compare characteristics between groups

- analyze relation of HDLox to other biomarkers

- predict AD, try to find a reasonably small model with and without HDLox

# Setup

## Packages

```{r}
pacman::p_load(
  conflicted, tidyverse, here, ggplot2, wrappedtools, car, broom,
  ggbeeswarm, readxl, ggsignif, flextable, DescTools, lme4, easystats,
  ggsignif, ggh4x, patchwork, GGally)

conflicts_prefer(
  dplyr::select,
  dplyr::filter,
  ggplot2::mean_cl_boot,
)
```

## Data preparation

Read in data.

```{r}
rawdata <- read_excel(
  path = here("Data/DatensatzHDLox.xlsx"),
  range = cell_cols("A:AB"),
)
head(rawdata)
```

Removing all rows, where column "HDLox, no unit" has NAs.

```{r}
rawdata <- rawdata |> filter(!is.na(rawdata[[names(rawdata)[5]]])) |>
  mutate(
    `Angiographically confirmed CAD (1=ja, 0=nein), 0 bedeudet angiogr. Ausgeschlossen` = 
      replace_na(`Angiographically confirmed CAD (1=ja, 0=nein), 0 bedeudet angiogr. Ausgeschlossen`, 0)
  )
head(rawdata)

```

Recoding 1/0 into factors with level yes/no. First find all columns where "0=nein" is in column header and change into factors.

```{r}
rawdata <- rawdata |>
  mutate(across(contains("0=nein"),
                ~ factor(.x, levels = c(0, 1),
                         labels = c("no", "yes"))))
head(rawdata)
```

Missed some cases, e.g. "Nein=0" or "0= nein"

```{r}
rawdata <- rawdata |>
  mutate(across(contains("0= nein"),
                ~ factor(.x, levels = c(0, 1),
                         labels = c("no", "yes")))) |>
  mutate(across(contains("Nein=0"),
                ~ factor(.x, levels = c(0, 1),
                         labels = c("no", "yes"))))
head(rawdata)
```

Recoding 1/2 into factors with level male/female. First find all columns where "1=male" is in column header and change into factors.

```{r}
rawdata <- rawdata |>
  mutate(across(contains("1=male"),
                ~ factor(.x, levels = c(0, 1),
                         labels = c("male", "female")))) |>
  mutate(across(contains("Adipo"),
                ~ factor(.x, levels = c(0, 1),
                         labels = c("no", "yes"))))
head(rawdata)
```

Casting Patienten-ID into factor

```{r}
rawdata <- rawdata |> mutate(
  `Patienten-ID` = factor(`Patienten-ID`))
```

Renaming columns using regex: `\\s*` for catching the zero or more white spaces before `\\(`, which in turn catches "(". `.*` will find everything after the opening parenthesis. The second regex finds all "in" as in "HDL in mg/dl" and replaces it with a whitespace. Third finds the units and places them inside \[\].

```{r}
old_names <- names(rawdata)
# str_view(names(rawdata), "\\s*\\(.*") # Used to to find regex

rawdata <- rawdata |>
  rename_with(~ str_remove(.x, "\\s*\\(.*")) |>
  rename_with(~ str_replace(.x, "\\s{1}in\\s{1}", " ")) |>
  rename_with(~ str_replace(.x, "\\s{1}(\\w+/\\w+)", " [\\1]"))

# Comparing old names with new names
tibble(old = old_names, new = names(rawdata)) |> print(n = Inf)

# Final name adjustment can be done manually
rawdata <- rawdata |>
  rename(
    ID = `Patienten-ID`,
    HDLox = `HDLox, no unit`,
    `LP(a)` = LP,
    `HbA1c [%]` = `HbA1c, %`,
    `Albumin [mg/g]` = `Albumin im Spoturin. [mg/gKrea]`,
    `eGFR [ml/min/1.73m2]` = `eGFR, [ml/min]/1,73m2`,
    `Diabetes mellitus` = `Diabetes mell.`,
    Birthyear = Geburtsdatum, 
    `Size [cm]` = Größe,
    `Weight [kg]` = Gewicht,
    IMT = `IMT average`) |>
  mutate(Smoking = factor(Smoking, levels = c(0, 1),
                          labels = c("no", "yes")))

for (row_i in seq(length(rawdata$`LP(a)`))) {
  rawdata$`LP(a)`[row_i] <- str_remove_all(rawdata$`LP(a)`[row_i], "<")
  rawdata$`LP(a)`[row_i] <- str_replace(rawdata$`LP(a)`[row_i], ",", ".")
}
rawdata <- rawdata |>
  mutate(`LP(a)` = as.numeric(`LP(a)`))

tibble(old = old_names, new = names(rawdata)) |> print(n = Inf)
rm(old_names)
```

Changing birthdays into birthyears.

```{r}
str_view(rawdata$Birthyear, "-\\d{2}-\\d{2}")
rawdata$Birthyear <- rawdata$Birthyear |> str_remove_all("-\\d{2}-\\d{2}")

rawdata <- rawdata |> mutate(
  `LP(a) [nmol/l]` = as.numeric(`LP(a)`),
  Birthyear = factor(Birthyear))
```

# Data Exploration

## Test for normality

This would be internally and wouldn't go out. It's my own exploration. Plotting continues data and check for normality, first using a graphical exploration.

```{r}
numvars <- ColSeeker(varclass = c("numeric"))
factvars <- ColSeeker(varclass = c("factor"),
                      exclude = c("ID", "Birthyear"))

plot_list <- list()
for (measurement_i in numvars$names) {
  plot_temp <- ggplot(rawdata, aes(x = .data[[measurement_i]])) +
    geom_density(fill = "lightblue", alpha = 0.5) +
    stat_theodensity(distri = "norm", linewidth = 0.5, color = "red") 
  
  plot_list[[measurement_i]] <- plot_temp
}
combined_plot <- wrap_plots(plot_list, ncol = 3)
combined_plot

rm(plot_temp)
rm(measurement_i)
rm(combined_plot)
rm(plot_list)
```

Unexpectedly, "Size" looks strange, turns out there was a bad number in the data

```{r}
rawdata$`Size [cm]`[275] <- rawdata$`Size [cm]`[275] / 100
rawdata$`Size [cm]` <- rawdata$`Size [cm]` * 100
rawdata$BMI[275] <- rawdata$`Weight [kg]`[275] / 
  ((rawdata$`Size [cm]`[275]/100)^2)
```

```{r}
plot_list <- list()
for (measurement_i in numvars$names) {
  plot_temp <- ggplot(rawdata, aes(x = .data[[measurement_i]])) +
    geom_density(fill = "lightblue", alpha = 0.5) +
    stat_theodensity(distri = "norm", linewidth = 0.5, color = "red") 
  
  plot_list[[measurement_i]] <- plot_temp
}

combined_plot <- wrap_plots(plot_list, ncol = 3)
combined_plot

rm(plot_temp)
rm(measurement_i)
rm(combined_plot)
rm(plot_list)
```

Albumin also does not look right, let's explore.

```{r}
ggplot(rawdata, aes(ID, `Albumin [mg/g]`)) +
  geom_point()
```

The plot show, that most values are relative small and there are a handfull of outliers.

Testing for normality with Kolmogorov-Smirnov-test against normal distribution. Significance indicates non normal distributed data. The p-value will be overly critical, as the sample size is very large.

```{r}
rawdata |> 
  summarise(
    across(.cols = where(is.numeric),
           .fns = list(
             pKS = ~ksnormal(.x,lillie = FALSE) |> 
               formatP(mark = T)),
           .names = "{.col}")) |>
  pivot_longer(cols = where(is.character), # contains("_")
               names_to = "Measure",
               values_to = "p-value (KS-test)") |>
  flextable() |> set_table_properties(layout = "autofit") |> 
  set_caption("Test for normality")
```

Test within healthy subjects

```{r}
rawdata |> 
  group_by(`Healthy subjects`) |>
  summarise(
    across(.cols = where(is.numeric),
           .fns = list(
             pKS = ~ksnormal(.x,lillie = FALSE) |> 
               formatP(mark = T)),
           .names = "{.col}"), 
    .groups = "drop") |>
  pivot_longer(cols = where(is.character), # contains("_")
               names_to = "Measure",
               values_to = "p-value (KS-test)") |>
  flextable() |> set_table_properties(layout = "autofit") |> 
  set_caption("Test for normality") |> merge_v(j = 1) |> hline(i = 17)
```

##Test for log-normality log transformation of data.

```{r}
log_rawdata <- rawdata |>
  mutate(
    across(.cols = where(is.numeric),
           .fns = \(x) log(x)
    ))

log_rawdata |>
  summarise(
    across(.cols = where(is.numeric),
           .fns = list(
             pKS = ~ksnormal(.x,lillie = FALSE) |> 
               formatP(mark = T)),
           .names = "{.col}")) |>
  pivot_longer(cols = where(is.character), # contains("_")
               names_to = "Measure",
               values_to = "p-value (KS-test)") |>
  flextable() |> set_table_properties(layout = "autofit") |>
  set_caption("Test for log-normality")

plot_list <- list()
for (measurement_i in numvars$names) {
  plot_temp <- ggplot(rawdata, aes(x = .data[[measurement_i]])) +
    geom_density(fill = "lightblue", alpha = 0.5) +
    stat_theodensity(distri = "norm", linewidth = 0.5, color = "red") 
  
  plot_list[[measurement_i]] <- plot_temp
}

combined_plot <- wrap_plots(plot_list, ncol = 3)
combined_plot

rm(plot_temp)
rm(measurement_i)
rm(combined_plot)
rm(plot_list)

rm(log_rawdata)
```

## Combined result table

```{r}
rawdata |>
  group_by(`Healthy subjects`) |>
  summarise(
    across(
      .cols = where(is.numeric),
      .fns = list(
        pKS = \(x) ksnormal(x, lillie = FALSE) |>
          formatP(mark = TRUE),
        pKS_log = \(x) ksnormal(log(x), lillie = FALSE)|>
          formatP(mark = TRUE)
      ),
      .names = "{.col}__{.fn}"
    ),
    .groups = "drop"
  ) |>
  pivot_longer(
    cols = -`Healthy subjects`, # contains("_")
    names_to = c("Measure", ".value"),
    names_sep = "__"
  ) |>
  flextable() |> set_table_properties(layout = "autofit") |>
  set_caption("Test for normality") |> merge_v(j = 1) |>
  hline(i = 17)
```

alternative by Andreas

```{r}
gausstest <-
  rawdata |>
  group_by(`Healthy subjects`) |>
  summarize(across(
    where(is.numeric),
    list(
      "norm" = \(x) ksnormal(x,
                             lillie = FALSE
      ),
      "lognorm" = \(x) ksnormal(log(x),
                                lillie = FALSE
      )
    )
  )) |>
  pivot_longer(-`Healthy subjects`,
               names_to = c("Variable", ".value"),
               names_sep = "_"
  ) |>
  arrange(Variable, `Healthy subjects`) |>
  pivot_wider(
    names_from = `Healthy subjects`,
    values_from = contains("norm"),
    names_glue = "healthy:{`Healthy subjects`} {.value}"
  ) |>
  mutate(across(
    contains("Healthy"),
    ~ formatP(.x,
              ndigits = 5, mark = TRUE
    )
  ))
flextable(gausstest) |>
  bg(~ str_detect(`healthy:no norm`, "n.s."),
     j = 2, bg = "lightgreen"
  ) |>
  bg(~ str_detect(`healthy:yes norm`, "n.s."),
     j = 3, bg = "lightgreen"
  ) |>
  bg(~ str_detect(`healthy:no lognorm`, "n.s."),
     j = 4, bg = "lightgreen"
  ) |>
  bg(~ str_detect(`healthy:yes lognorm`, "n.s."),
     j = 5, bg = "lightgreen"
  ) |>
  set_table_properties(
    width = 1,
    layout = "autofit"
  )
```

Decision regarding normality: Trust healthy data more If both normal with and without treat log-normal as normal Everything exempt HDLox can be treated as normal. Log-transform of columns. These should also treated as ordinal

## Defining gausvars/ordvars

```{r}
# Add tiny before log transform to avoid infinte values because of 0
rawdata <- rawdata |>
  mutate(
    IMT_ln = log(IMT + 0.001),
    `LDL [mg/dl]_ln` = log(`LDL [mg/dl]` + 0.001),
    `LP(a) [nmol/l]_ln` = log(`LP(a) [nmol/l]` + 0.001),
    PLA2_ln = log(PLA2 + 0.001),
    `Triglyceride [mmol/l]_ln` = log(`Triglyceride [mmol/l]` + 0.001),
    hsCRP_ln = log(hsCRP + 0.001),
    `HDL [mg/dl]_ln` = log(`HDL [mg/dl]` + 0.001), 
    `Cholesterin [mmol/l]_ln` = log(`Cholesterin [mmol/l]` + 0.001),
    `NT proBNP [pg/ml]_ln` = log(`NT proBNP [pg/ml]` + 0.001)
  )

gausvars <- ColSeeker(
  varclass = "numeric",
  namepattern = c("ln", "Size", "Weight", "BMI", "Cholesterin .mmol/l._ln", "IMT"),
  exclude = c("LP.a.")
)
gausvars <- gausvars$names
# gausvars[5] <- "IMT"

ordvars <- ColSeeker(
  varclass = "numeric",
  exclude = c("_ln"),
  namepattern = c("DL", "ri", "hsCRP", "PLA2", "NTpro", "Age", 
                  "Albumin", "HbA1c", "NTproBNP [pg/ml]", "eGFR", 
                  "LP(a) [nmol/l]", "NT proBNP [pg/ml]")
)
ordvars
names(rawdata)
ordvars <- ordvars$names
ordvars[12] <- "NT proBNP [pg/ml]"
ordvars[13] <- "LP(a)"
```

## Categorical Data Visualizations

```{r}
for (var_i in factvars$names) {
  abs_plot <-
    ggplot(rawdata, aes(x = .data[[var_i]], 
                        fill = `Peripheral/Carotid artery disease`)) +
    geom_bar()
  
  rel_plot <-
    ggplot(rawdata, aes(x = .data[[var_i]], 
                        fill = `Peripheral/Carotid artery disease`)) +
    geom_bar(position = "fill") +
    scale_y_continuous("frequency",
                       labels = scales::percent)
  
  print(abs_plot / rel_plot)
}

for (measurement_i in numvars$names) {
  plot_temp <- ggplot(rawdata, aes(
    x = `Peripheral/Carotid artery disease`,
    y = .data[[measurement_i]],
    color = `Peripheral/Carotid artery disease`)) +
    geom_boxplot(colour = "black") +
    geom_beeswarm(alpha = 0.5)
  
  print(plot_temp)
}

rm(abs_plot, rel_plot, plot_temp, measurement_i, var_i)
```

# Descriptive Statistics

```{r}
rawdata |> summarise(
  across(
    all_of(numvars$names),
    .fns = list(
      n = \(x) n(),
      mean = \(x) mean(x, na.rm = TRUE) |> roundR(textout = FALSE),
      median = \(x) median(x, na.rm = TRUE) |> roundR(textout = FALSE),
      SD = \(x) sd(x, na.rm = TRUE) |> roundR(textout = FALSE),
      SEM = \(x) SEM(x) |> roundR(textout = FALSE) 
    ))) |>
  pivot_longer(
    cols = everything(),
    names_to = c("Variable", ".value"),
    names_sep = "_"
  ) |>
  flextable() |>
  set_table_properties(width = 1, layout = "autofit") |> 
  set_caption("Overall descriptive statistics")
```

# Statistical Analysis

## Normal distributed variables

Comparing for differences between the gausvars between healthy and unhealthy subjects

```{r}
groupvars <- ColSeeker(namepattern = "Healthy")
plot_list <- list()
result_list <- list()
```

Analysis using wrapped tools and compare2numvars. Extraction of test is missing for summary table.

```{r}
for (measurement_i in seq(length(gausvars))) {
  result_temp <- compare2numvars(
    data = rawdata,
    dep_vars = gausvars[measurement_i],
    indep_var = groupvars$names,
    gaussian = TRUE,
    n = TRUE,
    mark = FALSE
  )
  result_list[[measurement_i]] <- result_temp
  
  plot_temp <- ggplot(
    rawdata,
    aes(x = .data[[groupvars$names]],
        y = .data[[gausvars[measurement_i]]],
        color = `Healthy subjects`)) +
    geom_beeswarm(alpha = 0.25) +
    stat_summary(fun.data = mean_sdl,
                 color = "black") +
    geom_signif(
      comparisons = list(c(1, 2)),
      annotations = paste("p", formatP(
        result_list[[measurement_i]]$p,
        pretext = T, mark = T)),
      color = "black") +
    ylab(paste(gausvars[measurement_i], "\nmean \u00b1 95% CI")) +
    theme(legend.position = "none") +
    scale_y_continuous(expand = expansion(mult = 0.2))
  
  plot_list[[measurement_i]] <- plot_temp
  print(plot_temp)
}
combined_result <- bind_rows(result_list)
combined_result |> flextable() |> 
  set_table_properties(layout = "autofit") |>
  bg(~ `p` < 0.05,
     j = 8, bg = "lightgreen"
  )

combined_plot <- wrap_plots(plot_list, ncol = 3, axes = "keep")
combined_plot

rm(plot_temp)
rm(measurement_i)
rm(combined_plot)
rm(plot_list)

```

Alternative

```{r}
plot_list <- list()
result_list <- list()

for (measurement_i in seq(length(gausvars))) {
  temp_x = rawdata[[gausvars[measurement_i]]][which(rawdata$`Healthy subjects` == "no")]
  temp_y = rawdata[[gausvars[measurement_i]]][which(rawdata$`Healthy subjects` == "yes")]
  
  result_temp <- t.test(
    x = temp_x,
    y = temp_y,
    var.equal = var.test(
      rawdata[[gausvars[measurement_i]]] ~ rawdata$`Healthy subjects`)$p.value > .05
  )
  
  result_temp_tibble <- tibble(
    Measure = gausvars[measurement_i],
    Method = result_temp$method,
    t = roundR(result_temp$statistic[[1]]),
    df = roundR(result_temp$parameter[[1]]),
    `p-value` = formatP(
      result_temp$p.value, mark = TRUE, pretext = TRUE)
  )
  
  result_list[[measurement_i]] <- result_temp_tibble
  
  plot_temp <- ggplot(
    rawdata,
    aes(x = .data[[groupvars$names]],
        y = .data[[gausvars[measurement_i]]],
        color = `Healthy subjects`)) +
    geom_beeswarm(alpha = 0.25) +
    stat_summary(fun.data = mean_cl_normal,
                 color = "black") +
    geom_signif(
      comparisons = list(c(1, 2)),
      annotations = paste("p", formatP(
        result_temp$p.value,
        pretext = T, mark = T)),
      color = "black") +
    ylab(paste(gausvars[measurement_i], "\nmean \u00b1 95% CI")) +
    theme(legend.position = "none") +
    scale_y_continuous(expand = expansion(mult = 0.2))
  
  plot_list[[measurement_i]] <- plot_temp
  print(plot_temp)
}
combined_result <- bind_rows(result_list)
combined_result |> flextable() |> 
  set_table_properties(layout = "autofit")

combined_plot <- wrap_plots(plot_list, ncol = 3, axes = "keep")
combined_plot

rm(temp_x, temp_y, plot_temp, measurement_i, combined_plot, plot_list)

```

## Ordinal variables

```{r}
ordvars
rawdata <- drop_na(rawdata)

plot_list <- list()
result_list <- list()

for (measurement_i in seq(length(ordvars))) {
  temp_x = rawdata[[ordvars[measurement_i]]][which(rawdata$`Healthy subjects` == "no")]
  temp_y = rawdata[[ordvars[measurement_i]]][which(rawdata$`Healthy subjects` == "yes")]
  
  result_temp <- wilcox.test(
    x = temp_x,
    y = temp_y
  )
  
  result_temp_tibble <- tibble(
    Measure = ordvars[measurement_i],
    Method = result_temp$method,
    w = roundR(result_temp$statistic[[1]]),
    `p-value` = formatP(
      result_temp$p.value, mark = TRUE, pretext = TRUE)
  )
  
  result_list[[measurement_i]] <- result_temp_tibble
  
  plot_temp <- ggplot(
    rawdata,
    aes(x = .data[[groupvars$names]],
        y = .data[[ordvars[measurement_i]]],
        color = `Healthy subjects`)) +
    geom_beeswarm(alpha = 0.25) +
    stat_summary(fun.data = mean_cl_normal,
                 color = "black") +
    geom_signif(
      comparisons = list(c(1, 2)),
      annotations = paste("p", formatP(
        result_temp$p.value,
        pretext = T, mark = T)),
      color = "black") +
    ylab(paste(ordvars[measurement_i], "\nmean \u00b1 95% CI")) +
    theme(legend.position = "none") +
    scale_y_continuous(expand = expansion(mult = c(0, 0.1)))
  
  plot_list[[measurement_i]] <- plot_temp
  print(plot_temp)
}
combined_result <- bind_rows(result_list)
combined_result |> flextable() |> 
  set_table_properties(layout = "autofit")


combined_plot <- wrap_plots(plot_list, ncol = 3, axes = "keep")
combined_plot

rm(temp_x, temp_y, plot_temp, measurement_i, combined_plot, plot_list,
   result_list, result_temp, result_temp_tibble, row_i, 
   combined_result)
```

# Analyzing the relationship between HDLox and other biomarkers

## Correlation

```{r}
corr_tibble <- cor(
  rawdata |> select(all_of(gausvars)),
  method = "pearson"
  )
ggpairs( cor(
  rawdata |> select(all_of(gausvars)),
  method = "pearson"
  ) )

# Which biomarkers are correlating with HDLox?
cortestR(
  rawdata |> select(all_of(ordvars)),
  method = "spear",
) |> 
  rownames_to_column("Reference") |>
  select(1:2) |>
  slice(-1) |>
  flextable() |> 
  set_table_properties(layout = "autofit")
```

## Linear Models

```{r}
rawdata <- drop_na(rawdata)

result_list <- list()

for (gaus_i in seq(gausvars)) {
  var_name <- gausvars[gaus_i]
  form_temp <- reformulate(
    termlabels = paste0("`", var_name, "`"), 
    response = "HDLox")
  
  plot_temp <- ggplot(
    rawdata,
    aes(x = .data[[gausvars[gaus_i]]],
        y = .data$HDLox)) +
    geom_point() +
    geom_smooth(method = "lm", color = "red") +
    geom_smooth(method = "loess", color = "blue")
  print(plot_temp)
  
  lm_out <- lm(form_temp, data = rawdata)
  anova_out <- anova(lm_out) |> tidy() |>
    mutate(
      p.value = formatP(p.value, ndigits = 5, 
                        mark = TRUE, pretext = TRUE)) |>
    add_column(measurement = gausvars[gaus_i] , .before = 1)
  
  result_list[[gaus_i]] <- anova_out
 }

combined_result <- bind_rows(result_list)
combined_result |> flextable() |> 
  set_table_properties(layout = "autofit") |> merge_v(j = 1)

rm(form_temp, gaus_i, plot_temp, var_name, result_list, lm_out, 
   anova_out, combined_result)
```

Adjusting for healthy/not healthy

```{r}
result_list <- list()

for (gaus_i in seq(gausvars)) {
  var_name <- gausvars[gaus_i]
  form_temp <- reformulate(
    termlabels = c(paste0("`", var_name, "`"), "`Healthy subjects`"), 
    response = "HDLox")
  
  lm_out <- lm(form_temp, data = rawdata)
  anova_out <- anova(lm_out) |> tidy() |>
    mutate(
      p.value = formatP(p.value, ndigits = 5, 
                        mark = TRUE, pretext = TRUE)) |>
    add_column(measurement = gausvars[gaus_i] , .before = 1)
  
  result_list[[gaus_i]] <- anova_out
}
combined_result <- bind_rows(result_list)
combined_result |> flextable() |> 
  set_table_properties(layout = "autofit") |> merge_v(j = 1) |>
  set_caption("Anova of Linear Models adjusted for healthy subjects")
```

# Building a multivariable additive model

Using those independent variables, which have shown significance in the univaribale model.

```{r}
indep_var <- c("`Weight [kg]`", "BMI", "`LDL [mg/dl]_ln`", 
               "`Triglyceride [mmol/l]_ln`", "hsCRP_ln", 
               "`HDL [mg/dl]_ln`", "`Cholesterin [mmol/l]_ln`")
```

# Modeling outcome of interest Artery Disease (AD)

As Artery Disease is a binary value, a generalized linear model with a logit link to a binomial distribution will be used.

## Univariable

```{r}
indep_var

rawdata <- rawdata |> rename(PAD = `Peripheral/Carotid artery disease`)
result_list <- ""
for (var_i in seq(length(indep_var))) {
  form_temp <- reformulate(indep_var[var_i], response = "PAD")
  glm_out <- glm(form_temp, family = "binomial", data = rawdata)
  result_list <- tidy(glm_out)
}

for (var_i in seq(length(ordvars))) {
  current_index <- var_i + length(gausvars)
  predvars[current_index] <- ordvars[var_i]
}
rawdata$`Peripheral/Carotid artery disease`
```

## Building a multivariate additive model

Building with significant measures from the univariable model ORs are difficult to interpret, for prediction univariable is better.

```{r}
uni_glmms <- tibble(predictor = NA_character_,
                    OR = NA_real_,
                    CI_low = NA_real_,
                    CI_high = NA_real_,
                    CIs = NA_character_,
                    p.value = NA_character_,
                    .rows = 0)
uni_anova <- tibble(variable = NA_character_,
                    p.value = NA_character_,
                    AIC = NA_real_,
                    BIC = NA_real_,
                    .rows = 0)

for (pred_i in seq(predvars)) {
  form <- reformulate(
    predvars[pred_i], 
    response = `Peripheral/Carotid artery disease`)
  uni_glmm <- glm(`Peripheral/Carotid artery disease` ~ `Size [cm]`, data = rawdata)
 }
```
